---
title: "Byonic Processing Notebook"
output: html_notebook
---
---
Libraries, constants & ion calculation functions
---
```{r}
# Define file path to the folder containing your data
directory <- c("C:/Users/your_file_path/")
setwd(directory)

# User defined variables
# TMT 6-plex ETD ion masses
TMT_ETD_mz_values <- c(114.127725, 115.124760, 116.134433, 117.131468, 118.141141, 119.138176)

# TMT 6-plex HCD ion masses
TMT_HCD_mz_values <- c(126.127726, 127.124761, 128.134436, 129.131471, 130.141145, 131.138180)

# TMT 6-plex isotopic corrections
# reference c(-2, -1, monoisotopic, +1, +2)
TMT126 <- c(0, 0, 1, 0.076, 0)
TMT127 <- c(0, 0, 1, 0.075, 0)
TMT128 <- c(0, 0.015, 1, 0.056, 0)
TMT129 <- c(0, 0.017, 1, 0.054, 0)
TMT130 <- c(0, 0.034, 1, 0.035, 0)
TMT131 <- c(0.007, 0.032, 1, 0.036, 0)
# Anal. Chem. 2012, 84, 16, 7188–7194
# reference calc: IC129 = IR129 - (IR128 * 0.017 + IR129 * 0.056)

# Define maximum number of modifications per peptide
max_mods <- 7

# Define the posterior error probability (PEP) limit for Byonic/Percolator results
PEP <- 0.05

# Define the False Discovery Rates (FDR) limit for Byonic results
#FDR <- 0.01

# Define the minimum delta mod score
min_delta_mod <- 10

# Define your fragment mass error tolerance in parts per million (ppm)
ppm_tolerance <- 15 

# Define the minimum intensity threshold at which isotopic checking of fragment ions is performed
#min_int_threshold <- 0.005

# Define whether to consider PSMs hits that are ambiguous based on detected ions
#unambiguous <- TRUE

# Define amino acid used for TMT mass tag entry in Byonic 
aa_tag <- "H"

library(parallel)
library(purrr)
library(mzR)
library(tidyr)
library(stringr)
library(dplyr)
library(xml2)
library(readr)
library(readxl)
library(writexl)
library(tidyverse) # includes ggplot2, for data visualisation. dplyr, for data manipulation.
library(RColorBrewer) # for a colourful plot
library(ggrepel) # for nice annotations

# Monoisotopic masses of amino acids
amino_acid_masses <- c(
  A = 71.03711, R = 156.10111, N = 114.04293, D = 115.02694,
  C = 103.00919, E = 129.04259, Q = 128.05858, G = 57.02146,
  H = 137.05891, I = 113.08406, L = 113.08406, K = 128.09496,
  M = 131.04049, F = 147.06841, P = 97.05276, S = 87.03203,
  T = 101.04768, W = 186.07931, Y = 163.06333, V = 99.06841
)

# Atomic masses for elements (Monoisotopic)
atomic_masses <- c(
  C = 12.0, c = 13.0033548378, H = 1.007825, h = 1.007276, N = 14.003074, n = 15.0001088982, O = 15.994915, P = 30.973762, S = 31.972071
)

# Adjustments for ion types - revised from disambiguator because Byonic accounts for TMT mass differently
ion_type_adjustments <- list(
  a = -(atomic_masses["C"] + atomic_masses["O"] + atomic_masses["H"]),
  b = 0, # [N]+[M]-H  ; N = mass of neutral N terminal group
  y = atomic_masses["H"] * (2) + atomic_masses["O"], # Neutral mass = [C]+[M]+H ; C = mass of neutral C-terminal group (OH)
  c = atomic_masses["H"] + (atomic_masses["H"] * (2) + atomic_masses["N"]), # [N]+[M]+NH2  ; N = mass of neutral N terminal group
  z = atomic_masses["O"] - atomic_masses["N"], # Neutral mass = [C]+[M]-NH2 ; C = mass of neutral C-terminal group (OH)
	# We're really printing Z-dot ions so we add an H to make it OH+[M]-NH2 +H = [M]+O-N
  Ion_AmmoniaLossMass = atomic_masses["H"] * 3 + atomic_masses["N"], # Define neutral loss masses
  Ion_WaterLossMass = atomic_masses["H"] * 2 + atomic_masses["O"] # Define neutral loss masses
)

# Define fn to calculate tolerance
calculate_tolerance <- function(mz, ppm) {
  return(as.numeric(mz) * ppm / 1e6)
}

# Function to calculate m/z for ions
calculate_ion_mz <- function(peptide_mass, charge) {
  # Adjust mass for ion type and add protons for charge
  mz <- (peptide_mass + atomic_masses["h"] * charge) / charge # h is proton
  return(mz)
}

calculate_fragment_mass_from_sequence <- function(sequence, modifications = NULL) {
  mass <- sum(sapply(strsplit(sequence, "")[[1]], function(aa) {
    if (aa %in% names(amino_acid_masses)) {
      return(amino_acid_masses[[aa]])
    } else {
      stop(paste("Undefined amino acid or modification:", aa))
    }
  }))
  
  proton_mass_subtracted <- 0
  if (!is.null(modifications) && nrow(modifications) > 0) {
    for (i in 1:nrow(modifications)) {
      mod <- modifications[i, ]
      if (mod$position <= nchar(sequence)) {
        mass <- mass + as.numeric(mod$variable)  # Add modification mass
        # Check for specific modification mass and subtract a proton's mass if applicable
        if (abs(mod$variable - 42.046950) < .Machine$double.eps^0.5) {
          mass <- mass - atomic_masses["h"]  # Subtract proton's mass for each occurrence
          proton_mass_subtracted <- proton_mass_subtracted + 1
        }
      }
    }
  }
  
  return(mass)
}

generate_ions_for_sequence <- function(sequence, modifications_df) {
  ions_list <- list()
  sequence_length <- nchar(sequence)
  
  for (charge in 1:5) { # Revised to +5 charge state max based on Byonic ion depiction
    a_ions <- numeric(sequence_length - 1)
    a_nl_O_ions <- numeric(sequence_length - 1)
    a_nl_N_ions <- numeric(sequence_length - 1)
    a_nl_Kme3_ions <- numeric(sequence_length - 1) # New vector for Kme3 neutral loss
    a_nl_pST_ions <- numeric(sequence_length - 1) # New vector for pST neutral loss
    a_nl_Ksu_ions <- numeric(sequence_length - 1) # New vector for Ksu neutral loss
    b_ions <- numeric(sequence_length - 1)
    b_nl_O_ions <- numeric(sequence_length - 1)
    b_nl_N_ions <- numeric(sequence_length - 1)
    b_nl_Kme3_ions <- numeric(sequence_length - 1) # New vector for Kme3 neutral loss
    b_nl_pST_ions <- numeric(sequence_length - 1) # New vector for pST neutral loss
    b_nl_Ksu_ions <- numeric(sequence_length - 1) # New vector for Ksu neutral losS
    y_ions <- numeric(sequence_length - 1)
    y_nl_O_ions <- numeric(sequence_length - 1)
    y_nl_N_ions <- numeric(sequence_length - 1)
    y_nl_Kme3_ions <- numeric(sequence_length - 1) # New vector for Kme3 neutral loss
    y_nl_pST_ions <- numeric(sequence_length - 1) # New vector for pST neutral loss
    y_nl_Ksu_ions <- numeric(sequence_length - 1) # New vector for Ksu neutral losS
    c_ions <- numeric(sequence_length - 1)
    c_nl_O_ions <- numeric(sequence_length - 1)
    c_nl_N_ions <- numeric(sequence_length - 1)
    c_nl_Kme3_ions <- numeric(sequence_length - 1) # New vector for Kme3 neutral loss
    c_nl_pST_ions <- numeric(sequence_length - 1) # New vector for pST neutral loss
    c_nl_Ksu_ions <- numeric(sequence_length - 1) # New vector for Ksu neutral losS
    z_ions <- numeric(sequence_length - 1)
    z_nl_O_ions <- numeric(sequence_length - 1)
    z_nl_N_ions <- numeric(sequence_length - 1)
    z_nl_Kme3_ions <- numeric(sequence_length - 1) # New vector for Kme3 neutral loss
    z_nl_pST_ions <- numeric(sequence_length - 1) # New vector for pST neutral loss
    z_nl_Ksu_ions <- numeric(sequence_length - 1) # New vector for Ksu neutral losS
    
    for (i in 1:(sequence_length - 1)) {
      a_seq <- substr(sequence, 1, i)
      b_seq <- substr(sequence, 1, i)
      y_seq <- substr(sequence, sequence_length - i + 1, sequence_length)
      c_seq <- substr(sequence, 1, i)
      z_seq <- substr(sequence, sequence_length - i + 1, sequence_length)
      
      # For b ions, use positions directly as they are from N-terminus
      a_mods <- modifications_df[modifications_df$position <= i,]
      b_mods <- modifications_df[modifications_df$position <= i,]
      c_mods <- modifications_df[modifications_df$position <= i,]
      
      # For y ions, adjust positions relative to the fragment's C-terminus
      y_mods <- modifications_df
      if (nrow(y_mods) > 0) {
        y_mods$position <- sequence_length - y_mods$position + 1
        y_mods <- y_mods[y_mods$position <= i,]
      }
      z_mods <- modifications_df
      if (nrow(z_mods) > 0) {
        z_mods$position <- sequence_length - z_mods$position + 1
        z_mods <- z_mods[z_mods$position <= i,]
      }

      # Check for specific modification by mass and calculate neutral loss
      if (any(abs(a_mods$variable - 42.046950) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Kme3_loss <- 59.073499  # Define this value
        a_mass_specific_mod_loss <- a_mass - specific_Kme3_loss
        a_nl_Kme3_ions[i] <- calculate_ion_mz(a_mass_specific_mod_loss, charge)
      }
      if (any(abs(b_mods$variable - 42.046950) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Kme3_loss <- 59.073499  # Define this value
        b_mass_specific_mod_loss <- b_mass - specific_Kme3_loss
        b_nl_Kme3_ions[i] <- calculate_ion_mz(b_mass_specific_mod_loss, charge)
      }
      if (any(abs(y_mods$variable - 42.046950) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Kme3_loss <- 59.073499  # Define this value
        y_mass_specific_mod_loss <- y_mass - specific_Kme3_loss
        y_nl_Kme3_ions[i] <- calculate_ion_mz(y_mass_specific_mod_loss, charge)
      }
      if (any(abs(c_mods$variable - 42.046950) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Kme3_loss <- 59.073499  # Define this value
        c_mass_specific_mod_loss <- c_mass - specific_Kme3_loss
        c_nl_Kme3_ions[i] <- calculate_ion_mz(c_mass_specific_mod_loss, charge)
      }
      if (any(abs(z_mods$variable - 42.046950) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Kme3_loss <- 59.073499  # Define this value
        z_mass_specific_mod_loss <- z_mass - specific_Kme3_loss
        z_nl_Kme3_ions[i] <- calculate_ion_mz(z_mass_specific_mod_loss, charge)
      }
      if (any(abs(a_mods$variable - 79.966331) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_pST_loss <- 97.976896  # Define this value
        a_mass_specific_mod_loss <- a_mass - specific_pST_loss
        a_nl_pST_ions[i] <- calculate_ion_mz(a_mass_specific_mod_loss, charge)
      }
      if (any(abs(b_mods$variable - 79.966331) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_pST_loss <- 97.976896  # Define this value
        b_mass_specific_mod_loss <- b_mass - specific_pST_loss
        b_nl_pST_ions[i] <- calculate_ion_mz(b_mass_specific_mod_loss, charge)
      }
      if (any(abs(y_mods$variable - 79.966331) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_pST_loss <- 97.976896  # Define this value
        y_mass_specific_mod_loss <- y_mass - specific_pST_loss
        y_nl_pST_ions[i] <- calculate_ion_mz(y_mass_specific_mod_loss, charge)
      }
      if (any(abs(c_mods$variable - 79.966331) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_pST_loss <- 97.976896  # Define this value
        c_mass_specific_mod_loss <- c_mass - specific_pST_loss
        c_nl_pST_ions[i] <- calculate_ion_mz(c_mass_specific_mod_loss, charge)
      }
      if (any(abs(z_mods$variable - 79.966331) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_pST_loss <- 97.976896  # Define this value
        z_mass_specific_mod_loss <- z_mass - specific_pST_loss
        z_nl_pST_ions[i] <- calculate_ion_mz(z_mass_specific_mod_loss, charge)
      }
      if (any(abs(a_mods$variable - 100.01604) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Ksu_loss <- 101.02387  # Define this value
        a_mass_specific_mod_loss <- a_mass - specific_Ksu_loss
        a_nl_Ksu_ions[i] <- calculate_ion_mz(a_mass_specific_mod_loss, charge)
      }
      if (any(abs(b_mods$variable - 100.01604) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Ksu_loss <- 101.02387  # Define this value
        b_mass_specific_mod_loss <- b_mass - specific_Ksu_loss
        b_nl_Ksu_ions[i] <- calculate_ion_mz(b_mass_specific_mod_loss, charge)
      }
      if (any(abs(y_mods$variable - 100.01604) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Ksu_loss <- 101.02387  # Define this value
        y_mass_specific_mod_loss <- y_mass - specific_Ksu_loss
        y_nl_Ksu_ions[i] <- calculate_ion_mz(y_mass_specific_mod_loss, charge)
      }
      if (any(abs(c_mods$variable - 100.01604) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Ksu_loss <- 101.02387  # Define this value
        c_mass_specific_mod_loss <- c_mass - specific_Ksu_loss
        c_nl_Ksu_ions[i] <- calculate_ion_mz(c_mass_specific_mod_loss, charge)
      }
      if (any(abs(z_mods$variable - 100.01604) < .Machine$double.eps^0.5)) {
        # Assuming a specific neutral loss for this modification
        specific_Ksu_loss <- 101.02387  # Define this value
        z_mass_specific_mod_loss <- z_mass - specific_Ksu_loss
        z_nl_Ksu_ions[i] <- calculate_ion_mz(z_mass_specific_mod_loss, charge)
      }
      
      # Calculate the mass of the b and y ion sequences
      a_mass <- calculate_fragment_mass_from_sequence(a_seq, a_mods) + ion_type_adjustments$a
      b_mass <- calculate_fragment_mass_from_sequence(b_seq, b_mods) + ion_type_adjustments$b
      y_mass <- calculate_fragment_mass_from_sequence(y_seq, y_mods) + ion_type_adjustments$y
      c_mass <- calculate_fragment_mass_from_sequence(c_seq, c_mods) + ion_type_adjustments$c
      z_mass <- calculate_fragment_mass_from_sequence(z_seq, z_mods) + ion_type_adjustments$z
      
      # Check for potential neutral losses and calculate accordingly
      if (grepl("[KRQN]", a_seq)) {
        a_mass_N_loss <- a_mass - ion_type_adjustments$Ion_AmmoniaLossMass
        a_nl_N_ions[i] <- calculate_ion_mz(a_mass_N_loss, charge)
      }
      if (grepl("[STED]", a_seq)) {
        a_mass_O_loss <- a_mass - ion_type_adjustments$Ion_WaterLossMass
        a_nl_O_ions[i] <- calculate_ion_mz(a_mass_O_loss, charge)
      }
      if (grepl("[KRQN]", b_seq)) {
        b_mass_N_loss <- b_mass - ion_type_adjustments$Ion_AmmoniaLossMass
        b_nl_N_ions[i] <- calculate_ion_mz(b_mass_N_loss, charge)
      }
      if (grepl("[STED]", b_seq)) {
        b_mass_O_loss <- b_mass - ion_type_adjustments$Ion_WaterLossMass
        b_nl_O_ions[i] <- calculate_ion_mz(b_mass_O_loss, charge)
      }
      # Check for potential neutral losses and calculate accordingly
      if (grepl("[KRQN]", y_seq)) {
        y_mass_N_loss <- y_mass - ion_type_adjustments$Ion_AmmoniaLossMass
        y_nl_N_ions[i] <- calculate_ion_mz(y_mass_N_loss, charge)
      }
      if (grepl("[STED]", y_seq)) {
        y_mass_O_loss <- y_mass - ion_type_adjustments$Ion_WaterLossMass
        y_nl_O_ions[i] <- calculate_ion_mz(y_mass_O_loss, charge)
      }
      # Check for potential neutral losses and calculate accordingly
      if (grepl("[KRQN]", c_seq)) {
        c_mass_N_loss <- c_mass - ion_type_adjustments$Ion_AmmoniaLossMass
        c_nl_N_ions[i] <- calculate_ion_mz(c_mass_N_loss, charge)
      }
      if (grepl("[STED]", c_seq)) {
        c_mass_O_loss <- c_mass - ion_type_adjustments$Ion_WaterLossMass
        c_nl_O_ions[i] <- calculate_ion_mz(c_mass_O_loss, charge)
      }
      # Check for potential neutral losses and calculate accordingly
      if (grepl("[KRQN]", z_seq)) {
        z_mass_N_loss <- z_mass - ion_type_adjustments$Ion_AmmoniaLossMass
        z_nl_N_ions[i] <- calculate_ion_mz(z_mass_N_loss, charge)
      }
      if (grepl("[STED]", z_seq)) {
        z_mass_O_loss <- z_mass - ion_type_adjustments$Ion_WaterLossMass
        z_nl_O_ions[i] <- calculate_ion_mz(z_mass_O_loss, charge)
      }
      
      a_ions[i] <- calculate_ion_mz(a_mass, charge)
      b_ions[i] <- calculate_ion_mz(b_mass, charge)
      y_ions[i] <- calculate_ion_mz(y_mass, charge)
      c_ions[i] <- calculate_ion_mz(c_mass, charge)
      z_ions[i] <- calculate_ion_mz(z_mass, charge)
    }
    
    ions_list[[paste("a_ions_charge", charge)]] <- a_ions
    ions_list[[paste("a_nl_O_ions_charge", charge)]] <- a_nl_O_ions
    ions_list[[paste("a_nl_N_ions_charge", charge)]] <- a_nl_N_ions
    ions_list[[paste("a_nl_Kme3_ions_charge", charge)]] <- a_nl_Kme3_ions
    ions_list[[paste("a_nl_pST_ions_charge", charge)]] <- a_nl_pST_ions
    ions_list[[paste("a_nl_Ksu_ions_charge", charge)]] <- a_nl_Ksu_ions
    ions_list[[paste("b_ions_charge", charge)]] <- b_ions
    ions_list[[paste("b_nl_O_ions_charge", charge)]] <- b_nl_O_ions
    ions_list[[paste("b_nl_N_ions_charge", charge)]] <- b_nl_N_ions
    ions_list[[paste("b_nl_Kme3_ions_charge", charge)]] <- b_nl_Kme3_ions
    ions_list[[paste("b_nl_pST_ions_charge", charge)]] <- b_nl_pST_ions
    ions_list[[paste("b_nl_Ksu_ions_charge", charge)]] <- b_nl_Ksu_ions
    ions_list[[paste("y_ions_charge", charge)]] <- y_ions
    ions_list[[paste("y_nl_O_ions_charge", charge)]] <- y_nl_O_ions
    ions_list[[paste("y_nl_N_ions_charge", charge)]] <- y_nl_N_ions
    ions_list[[paste("y_nl_Kme3_ions_charge", charge)]] <- y_nl_Kme3_ions
    ions_list[[paste("y_nl_pST_ions_charge", charge)]] <- y_nl_pST_ions
    ions_list[[paste("y_nl_Ksu_ions_charge", charge)]] <- y_nl_Ksu_ions
    ions_list[[paste("c_ions_charge", charge)]] <- c_ions
    ions_list[[paste("c_nl_O_ions_charge", charge)]] <- c_nl_O_ions
    ions_list[[paste("c_nl_N_ions_charge", charge)]] <- c_nl_N_ions
    ions_list[[paste("c_nl_Kme3_ions_charge", charge)]] <- c_nl_Kme3_ions
    ions_list[[paste("c_nl_pST_ions_charge", charge)]] <- c_nl_pST_ions
    ions_list[[paste("c_nl_Ksu_ions_charge", charge)]] <- c_nl_Ksu_ions
    ions_list[[paste("z_ions_charge", charge)]] <- z_ions
    ions_list[[paste("z_nl_O_ions_charge", charge)]] <- z_nl_O_ions
    ions_list[[paste("z_nl_N_ions_charge", charge)]] <- z_nl_N_ions
    ions_list[[paste("z_nl_Kme3_ions_charge", charge)]] <- z_nl_Kme3_ions
    ions_list[[paste("z_nl_pST_ions_charge", charge)]] <- z_nl_pST_ions
    ions_list[[paste("z_nl_Ksu_ions_charge", charge)]] <- z_nl_Ksu_ions
  }
  
  return(ions_list)
}

# Matching function
find_matches_with_intensity <- function(ions_df, scan_data, ppm_tolerance, sequence) {
  matches <- list()
  sequence_length <- nchar(sequence)
  max_intensity <- max(scan_data[, "intensity"], na.rm = TRUE)
  
  # Identify ion type columns in the dataframe
  ion_columns <- names(ions_df)
  
  for (ion_column in ion_columns) {
    # Determine the charge state based on the column name
    charge_state <- as.numeric(gsub(".*charge[ .](\\d+)$", "\\1", ion_column))
    ion_vector <- ions_df[[ion_column]]
    
    for (i in seq_along(ion_vector)) {
      ion_mass <- ion_vector[i]

      # Early skip conditions based on charge state and position in sequence
      if ((charge_state == 1 && i > (sequence_length / 4)) ||
          (charge_state == 2 && i > (sequence_length / 3)) ||
          (charge_state == 3 && i > (sequence_length / 3)) ||
          (charge_state == 4 && i < 4) ||
          (charge_state == 5 && i < (sequence_length / 3)) ||
          (charge_state == 6 && i < (sequence_length / 2))) {
        next
      }
      
      ion_tolerance <- calculate_tolerance(ion_mass, ppm_tolerance)
      
      # Find all potential matching indices
      all_matching_indices <- which(abs(scan_data[, "mz"] - ion_mass) <= ion_tolerance)

      if (length(all_matching_indices) > 0) {
        # Calculate distances for all matches
        distances <- abs(scan_data[all_matching_indices, "mz"] - ion_mass)
        # Select the index with the smallest distance
        closest_match_index <- all_matching_indices[which.min(distances)]

        # Calculate expected isotopic m/z and its tolerance
        expected_isotopic_mz <- scan_data[closest_match_index, "mz"] + (1 / charge_state)
        isotopic_tolerance <- calculate_tolerance(expected_isotopic_mz, ppm_tolerance)
        isotopic_peak_indices <- which(abs(scan_data[, "mz"] - expected_isotopic_mz) <= isotopic_tolerance)

        if (length(isotopic_peak_indices) > 0) {
          distances <- abs(scan_data[isotopic_peak_indices, "mz"] - expected_isotopic_mz)
          nearest_peak_index <- isotopic_peak_indices[which.min(distances)]

          if (scan_data[nearest_peak_index, "intensity"] <= scan_data[closest_match_index, "intensity"]) {
            # Construct a key for this match, including the column name and ion position
            match_key <- paste(ion_column,
                               "pos",
                               i,
                               "mz",
                               round(ion_mass, 4),
                               "charge",
                               charge_state,
                               sep = "_")
          # Store match with its intensity
            matches[[match_key]] <- list(mz = scan_data[closest_match_index, "mz"],
                                         intensity = scan_data[closest_match_index, "intensity"]
                                         )
          }
        }
      }
    }
  }

  return(matches)
}

# Key Points:
# Charge State Detection: The function now includes logic to determine the charge state from the column name. This uses a regular expression with gsub() to extract the charge state number from the column name pattern like "y_ions_charge.1". Adjust the pattern as necessary to match your dataframe's column naming convention.

# Biochemical Rule Application: Before proceeding with the match finding for each ion, it checks if the ion is in the +1 charge state and if its position is greater than half the length of the sequence. If both conditions are met, the next statement skips the rest of the current iteration, effectively filtering out these ions from the match list.

# Sequence Length Consideration: The function now requires the sequence parameter to calculate the sequence length and thereby apply the biochemical rule accurately.

# This approach allows you to integrate specific biochemical rules into the process of identifying and retaining meaningful matches, enhancing the biological relevance of your analysis results.

stored_parameters <- list(
  max_mods = max_mods,
  PEP.2D = PEP,
#  FDR.2D = FDR,
  min_delta_mod = min_delta_mod,
  ppm_tolerance = ppm_tolerance,
  aa_tag = aa_tag
#  min_int_threshold = min_int_threshold,
#  unambiguous = unambiguous
)

save_name <- function(stored_parameters) {
  # Generate a file name based on the parameters
  file_name <- paste0(
    "Byonic_",
    "max_mods_", stored_parameters$max_mods, "_",
    "PEP.2D_", stored_parameters$PEP, "_",
#    "FDR.2D_", stored_parameters$FDR, "_",
    "min_delta_mod_", stored_parameters$min_delta_mod, "_",
    "ppm_tolerance_", stored_parameters$ppm_tolerance, "_",
    stored_parameters$aa_tag, "-tag"
#    "min_int_threshold_", stored_parameters$min_int_threshold, "_",
#    "unambiguous_", stored_parameters$unambiguous
  )
  return(file_name)
}

save_name_suffix <- save_name(stored_parameters)
save_name_RData_suffix <- paste0(save_name(stored_parameters), ".RData")
save_name_csv_suffix <- paste0(save_name(stored_parameters), ".csv")
```

Read in Byonic data

```{r}
setwd(directory)

# Export Byonic search results as a csv
# Read in Byonic search results from csv
Byonic_df <- read.csv("Byonic_results.csv")

Byonic_df_backup <- Byonic_df
Byonic_df <- Byonic_df_backup

Byonic_df <- Byonic_df %>%
#  mutate(FDR.2D = as.numeric(FDR.2D)) %>%
#  filter(FDR.2D <= FDR) %>%
  mutate(PEP.2D = as.numeric(PEP.2D),
         Delta.Mod..Score = as.numeric(Delta.Mod..Score),
         Obs..Frag..Sum = as.integer(Obs..Frag..Sum),
         Scan.Number.s...Posit. = as.integer(Scan.Number.s...Posit.)) %>%
  filter(PEP.2D <= PEP) %>%
  filter(Delta.Mod..Score >= min_delta_mod) #%>%
#  extract(col = Scan.., into = "scan_number", regex = "scan=(\\d+)", remove = FALSE) %>%
#  mutate(scan_number = as.integer(scan_number))

print(unique(Byonic_df$Sequence..unformatted.))

Byonic_df <- Byonic_df %>%
  filter(!(str_detect(Scan.Time.s...Posit., ";"))) %>%
  filter(
    str_detect(Sequence..unformatted., "ARTKQTARKSTGGKAPRKQLATKAARKSAPATGGGH") |
    str_detect(Sequence..unformatted., "ARTKQTARKSTGGKAPRKQLATKAARKSAPSTGGGH")
  ) %>%
  select("Row.", "PID", "Protein.name", "Sequence..unformatted.", "Sequence", "Score", "Delta.Score", "Delta.Mod..Score", "XIC.area.summed", "XIC.AUC", "XIC.Start", "XIC.End", "Obs..Frag..Sum", "Scan.Number.s...Posit.", "Mod..AAs", "Mod..Names", "Mod..Summary", "Var..Pos..Protein", "PEP.2D", "PEP.1D", "Delta.Mod..Score", "Row.type", "Score.Rank", "MS2.Search.Alias.name", "Byonic.Comment", "Protein.alias.name", "Fixed.Mod..Summary", "MS.MS.Query", "MS.Alias.name") %>%
  mutate(PID = as.integer(PID))

# Combine Mod..Summary and Fixed.Mod..Summary into a new column 'modifications'
Byonic_df <- Byonic_df %>%
  mutate(modifications = case_when(
    Mod..Summary == "" & Fixed.Mod..Summary == "" ~ "",
    Mod..Summary == "" ~ Fixed.Mod..Summary,
    Fixed.Mod..Summary == "" ~ Mod..Summary,
    TRUE ~ paste(Mod..Summary, Fixed.Mod..Summary, sep = "; ")
  )) %>%
  mutate(modifications = str_replace_all(modifications, "Sortag2", "37"))

# Define a function to parse position and mass values and create a dataframe
parse_position_mass <- function(mod_str) {
  mod_list <- str_split(mod_str, ";\\s*")[[1]]
  mods_df <- map_df(mod_list, ~ {
    if (.x != "") {
      # Extract only the numeric part of the position
      position <- as.integer(str_extract(.x, "\\d+"))
      variable <- as.numeric(str_extract(.x, "(?<=/)[\\d\\.]+"))
      tibble(position = position, variable = variable)
    } else {
      tibble(position = NA_integer_, variable = NA_real_)
    }
  })
  
  # Check if both positions 36 and 37 exist in the dataframe
  if(all(c(36, 37) %in% mods_df$position)) {
    # Calculate the sum of the masses for positions 36 and 37
    sum_mass <- sum(mods_df$variable[mods_df$position %in% c(36, 37)], na.rm = TRUE)
    
    # Remove the original rows for position 37
    mods_df <- mods_df[!mods_df$position %in% c(37), ]
    
    # Update the mass at position 36 to the new sum
    if(any(mods_df$position == 36)) {
      mods_df <- mods_df %>%
        mutate(variable = if_else(position == 36, sum_mass, variable))
      # Change the position of this updated mass from 36 to 37
      mods_df <- mods_df %>%
        mutate(position = if_else(position == 36, 36, position))
    } else {
      # If position 36 doesn't exist, just add a new row for position 37
      mods_df <- bind_rows(mods_df, tibble(position = 36, variable = sum_mass))
    }
  }
  
  return(mods_df)
}

Byonic_df <- Byonic_df %>%
  rowwise() %>%
  mutate(modifications_df = list(parse_position_mass(modifications))) %>%
  ungroup()

# If multiple experiments were analyzed together in Byonic, filter them for individual analysis here
# Filtering by experiment name
experiment1_df <- Byonic_df %>% filter(MS.Alias.name %in% c("your_ms_experiment_1"))

# Filtering by experiment name
experiment2_df <- Byonic_df %>% filter(MS.Alias.name %in% c("your_ms_experiment_1"))

# Filtering by experiment name
experiment3_df <- Byonic_df %>% filter(MS.Alias.name %in% c("your_ms_experiment_1"))                

# Generic list function
#Byonic_list <- by(Byonic_df, 1:nrow(Byonic_df), function(row) {
#  list(scan_number = row$Scan.Number.s...Posit.,
#       sequence = row$Sequence..unformatted.,
#       PID = row$PID,
#       modifications_df = row$modifications_df)  # Assuming 'mz' is correctly structured for all workers
#})

experiment1_list <- by(experiment1_df, 1:nrow(experiment1_df), function(row) {
  list(scan_number = row$Scan.Number.s...Posit.,
       sequence = row$Sequence..unformatted.,
       PID = row$PID,
       modifications_df = row$modifications_df)  # Assuming 'mz' is correctly structured for all workers
})

experiment2_list <- by(experiment2_df, 1:nrow(experiment2_df), function(row) {
  list(scan_number = row$Scan.Number.s...Posit.,
       sequence = row$Sequence..unformatted.,
       PID = row$PID,
       modifications_df = row$modifications_df)  # Assuming 'mz' is correctly structured for all workers
})

experiment3_list <- by(experiment3_df, 1:nrow(experiment3_df), function(row) {
  list(scan_number = row$Scan.Number.s...Posit.,
       sequence = row$Sequence..unformatted.,
       PID = row$PID,
       modifications_df = row$modifications_df)  # Assuming 'mz' is correctly structured for all workers
})

experiment1_scan_list <- as.list(c(experiment1_df$Scan.Number.s...Posit.))

experiment2_scan_list <- as.list(c(experiment2_df$Scan.Number.s...Posit.))

experiment3_scan_list <- as.list(c(experiment3_df$Scan.Number.s...Posit.))

# Create list of mz intensity values
process_list <- function(input_list) {
  # Initialize list to store values
  empty_list <- list()

  # Iterate through Scan_list using retrieved scan_number values
  for(i in seq_along(input_list)) {
    scan_number <- as.character(input_list[[i]]$scan_number)
    PID <- as.character(input_list[[i]]$PID)
    scan_data <- input_list[[i]]
    empty_list[[as.character(scan_number)]][[PID]] <- scan_data
  }
  
  # Explicitly return the list
  return(empty_list)
}

experiment1_list <- process_list(experiment1_list)

experiment2_list <- process_list(experiment2_list)

experiment3_list <- process_list(experiment3_list)

```

Read in mass spec data

```{r}
setwd(directory)

# Function to process scans from thermo .RAW data converted to mzML.gz
# Create list of mz intensity values
process_scans <- function(Scan_list, mz) {
  # Initialize list to store values
  mz_list <- list()

  # Iterate through Scan_list using retrieved scan_number values
  for(i in seq_along(Scan_list)) {
    scan_number <- as.integer(Scan_list[[i]])
    scan_data <- as.data.frame(peaks(mz, scan = scan_number))
    mz_list[[as.character(scan_number)]] <- scan_data
  }
  
  # Explicitly return the list
  return(mz_list)
}

experiment1_mz <- openMSfile("C:/path_to_your_data/file1.mzML.gz")

experiment1_mz_list <- process_scans(experiment1_scan_list, experiment1_mz)

experiment2_mz <- openMSfile("C:/path_to_your_data/file2.mzML.gz")

experiment2_mz_list <- process_scans(experiment2_scan_list, experiment2_mz)

experiment3_mz <- openMSfile("C:/path_to_your_data/file3.mzML.gz")

experiment3_mz_list <- process_scans(experiment3_scan_list, experiment3_mz)
```

Find ion matches

```{r}
# Function to find ions in the raw data that match the PSM called by Byonic
list_match <- function(list1_chunk, list2) {
  matches_data <- list()  # Initialize outside the loops
  
  # Iterate through the chunk of list1 using names for direct access
  for (scan_number in names(list1_chunk)) {
    # Direct access to each scan_number's data in list1
    scan_data_list <- list1_chunk[[scan_number]]
    
    # Iterate through hits within each scan_number
    for (PID in names(scan_data_list)) {
      hit_data <- scan_data_list[[PID]]
      sequence <- hit_data[["sequence"]]
      modifications_df <- as.data.frame(hit_data[["modifications_df"]])  
  
      # Retrieve m/z and intensity data from list2 using scan_number
      scan_data <- list2[[scan_number]]
      
      # Generate ions for the sequence with its modifications
      ions_df <- generate_ions_for_sequence(sequence, modifications_df)
      
      # Find matching ions in the mass spectrometry data
      matches_df <- find_matches_with_intensity(ions_df, scan_data, ppm_tolerance, sequence)
    
      # Enrich matches_df with additional data
      matches_df$peptide_sequence <- sequence
      modifications_str <- apply(modifications_df, 1, function(x) paste(x['position'], x['variable'], sep=":"))
      matches_df$modifications <- paste(modifications_str, collapse="; ")
      
      # Store the match data indexed by scan_number and PID
      if (!exists(scan_number, matches_data)) {
        matches_data[[scan_number]] <- list()
      }
      matches_data[[scan_number]][[PID]] <- matches_df
    }
  }
  
  return(matches_data)
}

Byonic_list <- experiment1_list
mz_list <- experiment1_mz_list

# 1. Prepare List Chunks
numCores <- detectCores() / 2  # Save some cores for system tasks
chunks <- split(Byonic_list, cut(seq_along(Byonic_list), breaks = numCores, labels = FALSE))

# 2. Execute in Parallel
cl <- makeCluster(numCores)
on.exit(stopCluster(cl), add = TRUE)

# 3. Export necessary objects and functions to each cluster node
clusterExport(cl, varlist = c("find_matches_with_intensity", 
								              "generate_ions_for_sequence",
              								"calculate_tolerance",
              								"calculate_ion_mz",
              								"calculate_fragment_mass_from_sequence",
								              "ppm_tolerance",
								              "amino_acid_masses",
								              "atomic_masses",
								              "ion_type_adjustments",
								              "mz_list",
								              "list_match"))
clusterEvalQ(cl, {
  library(dplyr)  
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Execute the matching function in parallel
match_results_chunks <- parLapply(cl, chunks, function(chunk) {
  list_match(chunk, mz_list)
})

# Combine the lists from each chunk into a single list of matches
matches_list <- do.call(c, match_results_chunks)

stopCluster(cl)

# Correctly extracting the scan_number values assuming they are the part after the dot
corrected_names <- sapply(names(matches_list), function(x) {
    # Splitting the name based on the dot
    parts <- strsplit(x, "\\.")[[1]]
    
    # If there's a part after the dot, return it, otherwise return the original part
    if (length(parts) > 1) {
        return(parts[2])  # Return the part after the dot
    } else {
        return(x)  # If there's no dot, return the original name
    }
})

# Assign the corrected names back to the list
names(matches_list) <- corrected_names
```

Convert list to dataframe

```{r}
# 1. Prepare Data Chunks
numCores <- detectCores() / 2  # Save some cores for system tasks
chunks <- split(matches_list, cut(seq_along(matches_list), breaks = numCores, labels = FALSE))

total_chunk_length <- sum(sapply(chunks, length))
original_length <- length(matches_list)
if (total_chunk_length != original_length) {
  stop("Chunk lengths do not sum to the original length.")
}

# 2. Define a Function for Processing Chunks
process_chunk <- function(chunk) {
  # Initialize an empty dataframe for this chunk
  matches_list_df_chunk <- data.frame(
    scan_number = integer(),
    PID = integer(),
    ion_type = character(),
    ion_position = integer(),
    ion_charge = integer(),
    mz = numeric(),
    intensity = numeric(),
    peptide = character(),
    modifications = character(),
    stringsAsFactors = FALSE
  )

  # Initialize an empty list to collect rows for this chunk
  rows_to_add <- list()
  
  # Loop through each scan_number in the chunk
  for (scan_number in names(chunk)) {
    for (PID in names(chunk[[scan_number]])) {
      hit_data <- chunk[[scan_number]][[PID]]
      
      # Extract peptide_sequence and modifications for this scan_number and PID
      peptide_sequence <- hit_data[["peptide_sequence"]]
      modifications <- hit_data[["modifications"]]
    
      # Ensure ion_keys exclude metadata keys
      ion_keys = setdiff(names(hit_data), c("peptide_sequence", "modifications"))
  
      for (ion_key in ion_keys) { # ion_keys defined by setdiff() excluding metadata keys
        ion_data <- hit_data[[ion_key]]
  
       # Extract ion_type and ion_position from ion_key
        ion_type <- sub("^(.*?)_ions.*$", "\\1", ion_key) # This might need adjustment
        parts <- strsplit(ion_key, "_pos_")[[1]]
        ion_position <- as.integer(strsplit(parts[2], "_")[[1]][1])
  
        # Extract ion_charge from ion_key
        charge_pattern <- "_charge[ .]?([0-9]+)_"
        if(length(gregexpr(charge_pattern, ion_key)[[1]]) > 0) { # Ensure a match is found
          matches <- regmatches(ion_key, gregexpr(charge_pattern, ion_key))
          ion_charge <- as.integer(sub(charge_pattern, "\\1", matches[[1]]))
        } else {
          ion_charge <- NA # Assign NA if no charge information is extracted
        }
      
        # Collect the new row to add it to the list
        rows_to_add[[length(rows_to_add) + 1]] <- data.frame(
          scan_number = as.integer(scan_number),
          PID = as.integer(PID),
          ion_type = ion_type,
          ion_position = ion_position,
          ion_charge = ion_charge,
          mz = ion_data$mz,
          intensity = ion_data$intensity,
          peptide = peptide_sequence,
          modifications = modifications,
          stringsAsFactors = FALSE
        )
      }
    }
  }
  
  # After collecting all rows, combine them into a single dataframe
  matches_list_df_chunk <- do.call(rbind, rows_to_add)
}

# 3. Execute in Parallel
cl <- makeCluster(numCores)
on.exit(stopCluster(cl), add = TRUE)

# Assuming 'process_chunk' and necessary libraries are already exported to the cluster
match_results_df_chunks <- parLapply(cl, chunks, process_chunk)

# Combine the dataframes from each chunk
matches_data_df <- do.call(rbind, match_results_df_chunks)

matches_data_scans <- c(matches_data_df$scan_number)
Byonic_scans <- c(as.integer(names(Byonic_list)))

# Example for a single missing scan_number
missing_scan_numbers <- setdiff(Byonic_scans, matches_data_scans)
for(scan_number in missing_scan_numbers) {
  if(!is.null(Byonic_list[[as.character(scan_number)]])) {
    cat("Found in Byonic_list: ", scan_number, "\n")
  }
  # Additional checks can be applied similarly for intermediate or processing stages
}

#Omissions appear to be PSMs with no matched ions under the current criteria
```

Step-wise disambiguation
Step 1: split up modifications data

```{r}
# Assuming matches_data_df_w_mods already exists and has a 'modifications' column

# Define fill_modifications_direct function to work directly on the dataframe
fill_modifications_direct <- function(df, max_mods) {
  for (row in 1:nrow(df)) {
    modification_str <- df$modifications[row]
    if (is.na(modification_str)) next # Skip if NA
    
    mod_parts <- strsplit(modification_str, ";")[[1]]
    mods <- lapply(mod_parts, function(x) strsplit(x, ":")[[1]])
    
    for (i in 1:length(mods)) {
      if (i > max_mods) break # Do not exceed the max number of modifications
      mod_pair <- mods[[i]]
      df[row, paste0("mod_pos_", i)] <- as.numeric(mod_pair[1])
      df[row, paste0("mod_mass_", i)] <- as.numeric(mod_pair[2])
    }
  }
  return(df)
}

# Parallelization

# Step 1: Group your dataframe by scan_number
grouped_df_list <- matches_data_df %>% 
  group_by(scan_number) %>%
  group_split()

# Step 2: Prepare to use parallel processing
numCores <- detectCores() / 2  # Reserve two cores for system tasks
cl <- makeCluster(numCores)
on.exit(stopCluster(cl), add = TRUE)

# Step 3: Export necessary objects and functions to each cluster node
clusterExport(cl, varlist = c("fill_modifications_direct", "max_mods"))
clusterEvalQ(cl, {
  library(dplyr)  # Assuming dplyr is used in your functions
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Execute the task in parallel
fill_mods_results <- parLapply(cl, grouped_df_list, function(group) {
  fill_modifications_direct(group, max_mods)
})

# Combine the results from all cores
# The method to combine results depends on the structure of your output
matches_data_df_w_mods <- bind_rows(fill_mods_results)

add_column_if_not_exists <- function(df, column_name) {
  # Check if the column does not exist
  if (!column_name %in% names(df)) {
    # Create the column with NA values
    df[[column_name]] <- NA
  }
  # Return the modified dataframe
  return(df)
}

matches_data_df_w_mods <- add_column_if_not_exists(matches_data_df_w_mods, "mod_pos_6") 
matches_data_df_w_mods <- add_column_if_not_exists(matches_data_df_w_mods, "mod_mass_6")
matches_data_df_w_mods <- add_column_if_not_exists(matches_data_df_w_mods, "mod_pos_7") 
matches_data_df_w_mods <- add_column_if_not_exists(matches_data_df_w_mods, "mod_mass_7")

```

Step-wise disambiguation
Step 2: find nearest_before and nearest_after by modification type

```{r}
# HELPER functions

# Modified find_nearest_KR function to operate on dataframe columns
find_nearest_KR_df <- function(peptide, mod_position) {
  # Initialize vectors to store results
  nearest_before <- numeric(length(peptide))
  nearest_after <- numeric(length(peptide))
  
  for (i in seq_along(peptide)) {
    aa_sequence <- unlist(strsplit(peptide[i], ""))
    kr_positions <- which(aa_sequence %in% c("K", "R"))
    nearest_before[i] <- ifelse(any(kr_positions < mod_position[i]), max(kr_positions[kr_positions < mod_position[i]]), NA)
    nearest_after[i] <- ifelse(any(kr_positions > mod_position[i]), min(kr_positions[kr_positions > mod_position[i]]), NA)
  }
  
  # Return a dataframe with the calculated positions
  return(data.frame(nearest_before = nearest_before, nearest_after = nearest_after))
}

# Modified find_nearest_K function to operate on dataframe columns
find_nearest_K_df <- function(peptide, mod_position) {
  # Initialize vectors to store results
  nearest_before <- numeric(length(peptide))
  nearest_after <- numeric(length(peptide))
  
  for (i in seq_along(peptide)) {
    aa_sequence <- unlist(strsplit(peptide[i], ""))
    kr_positions <- which(aa_sequence %in% c("K"))
    nearest_before[i] <- ifelse(any(kr_positions < mod_position[i]), max(kr_positions[kr_positions < mod_position[i]]), NA)
    nearest_after[i] <- ifelse(any(kr_positions > mod_position[i]), min(kr_positions[kr_positions > mod_position[i]]), NA)
  }
  
  # Return a dataframe with the calculated positions
  return(data.frame(nearest_before = nearest_before, nearest_after = nearest_after))
}

# Modified find_nearest_ST function to operate on dataframe columns
find_nearest_ST_df <- function(peptide, mod_position) {
  # Initialize vectors to store results
  nearest_before <- numeric(length(peptide))
  nearest_after <- numeric(length(peptide))
  
  for (i in seq_along(peptide)) {
    aa_sequence <- unlist(strsplit(peptide[i], ""))
    kr_positions <- which(aa_sequence %in% c("S", "T"))
    nearest_before[i] <- ifelse(any(kr_positions < mod_position[i]), max(kr_positions[kr_positions < mod_position[i]]), NA)
    nearest_after[i] <- ifelse(any(kr_positions > mod_position[i]), min(kr_positions[kr_positions > mod_position[i]]), NA)
  }
  
  # Return a dataframe with the calculated positions
  return(data.frame(nearest_before = nearest_before, nearest_after = nearest_after))
}

# Assuming max_mods is already defined as the maximum number of modifications you expect per peptide
# For each modification, you want to add nearest_before_mod_n, nearest_after_mod_n, and nearest_fn

# Add empty columns for nearest_before_mod_n, nearest_after_mod_n, and nearest_fn for each modification
for(i in 1:max_mods) {
  matches_data_df_w_mods[paste0("nearest_before_mod_", i)] <- NA_real_
  matches_data_df_w_mods[paste0("nearest_after_mod_", i)] <- NA_real_
  matches_data_df_w_mods[paste0("nearest_fn_mod_", i)] <- NA_character_
}

# Assuming you have a dataframe `matches_data_df_w_mods` with peptides, modification positions, and masses

# Update the function to directly modify the dataframe
apply_nearest_function_directly <- function(df) {
  # Iterate over each row of the dataframe
  for (i in 1:nrow(df)) {
    # Extract current row's peptide, mod positions, and mod masses
    peptide <- df$peptide[i]
    mod_positions <- sapply(1:max_mods, function(n) df[[paste0("mod_pos_", n)]][i])
    mod_masses <- sapply(1:max_mods, function(n) df[[paste0("mod_mass_", n)]][i])

    # Initialize containers for the results
    nearest_befores <- rep(NA, max_mods)
    nearest_afters <- rep(NA, max_mods)
    nearest_fns <- rep(NA, max_mods)
    
    # Process each modification
    for (j in 1:max_mods) {
      mod_position <- mod_positions[j]
      mod_mass <- mod_masses[j]
      # Skip processing if mod_position or mod_mass is NA
      if (is.na(mod_position) || is.na(mod_mass)) next
      
      # Decide which nearest function to use based on mod_mass
      if (mod_mass %in% c(14.015650, 28.031300)) {
        result <- find_nearest_KR_df(peptide, mod_position)
        nearest_fn <- "KR" # updated to remove trimethylation from KR check
      } else if (mod_mass %in% c(79.966331)) {
        result <- find_nearest_ST_df(peptide, mod_position)
        nearest_fn <- "ST"
      } else if (mod_mass %in% c(42.010565, 42.046950, 56.026215, 72.02113, 86.036779, 100.01604)) {
        result <- find_nearest_K_df(peptide, mod_position)
        nearest_fn <- "K" # updated to add trimethylation given specificity for K
      } else {
        result <- list(nearest_before = NA, nearest_after = NA)
        nearest_fn <- "NA"
      }
      
      # Store the results
      nearest_befores[j] <- result$nearest_before
      nearest_afters[j] <- result$nearest_after
      nearest_fns[j] <- nearest_fn
    }

    # Assign the results back to the dataframe
    for (j in 1:max_mods) {
      df[[paste0("nearest_before_mod_", j)]][i] <- nearest_befores[j]
      df[[paste0("nearest_after_mod_", j)]][i] <- nearest_afters[j]
      df[[paste0("nearest_fn_mod_", j)]][i] <- nearest_fns[j]
    }
  }
  
  return(df)
}

# Apply the function to your dataframe
# Parallelization

library(parallel)
library(dplyr)
library(tidyr)
library(stringr)

# Step 1: Group your dataframe by scan_number
grouped_df_list <- matches_data_df_w_mods %>% 
  group_by(scan_number) %>%
  group_split()

# Step 2: Prepare to use parallel processing
numCores <- detectCores() / 2  # Reserve two cores for system tasks
cl <- makeCluster(numCores)
on.exit(stopCluster(cl), add = TRUE)

# Step 3: Export necessary objects and functions to each cluster node
clusterExport(cl, varlist = c("find_nearest_KR_df", "find_nearest_K_df", "find_nearest_ST_df", "max_mods", "apply_nearest_function_directly"))
clusterEvalQ(cl, {
  library(dplyr)  # Assuming dplyr is used in your functions
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Execute the task in parallel
find_nearest_results <- parLapply(cl, grouped_df_list, function(group) {
  apply_nearest_function_directly(group)
})

# Combine the results from all cores
# The method to combine results depends on the structure of your output
matches_data_df_w_mods_updated <- bind_rows(find_nearest_results)
```

Step-wise disambiguation
Step 3: use ion identities to determine whether fragments fall between nearest_before and nearest_after

```{r}
# Helper function

#First, we need a function that assembles a_b_c_ions_df and y_z_ions_df for each scan_number and PID combination from matches_data_df_w_mods_updated. This will involve filtering the dataframe for each ion type and converting y and z ion positions.
assemble_ions_df <- function(df, target_scan_number, target_PID) {
  a_b_c_ions_df <- dplyr::filter(df,
                               scan_number == as.numeric(target_scan_number),
                               PID == as.numeric(target_PID),
                               ion_type %in% c("a", "b", "c", "a_nl_O", "b_nl_O", "c_nl_O", "a_nl_N", "b_nl_N", "c_nl_N", "a_nl_Kme3", "b_nl_Kme3", "c_nl_Kme3", "a_nl_pST", "b_nl_pST", "c_nl_pST", "a_nl_Ksu", "b_nl_Ksu", "c_nl_Ksu"))

  y_z_ions_df <- dplyr::filter(df,
                               scan_number == as.numeric(target_scan_number),
                               PID == as.numeric(target_PID),
                               ion_type %in% c("y", "z", "y_nl_O", "z_nl_O", "y_nl_N", "z_nl_N", "y_nl_Kme3", "z_nl_Kme3", "y_nl_pST", "z_nl_pST", "y_nl_Ksu", "z_nl_Ksu")) %>%
    mutate(ion_position = nchar(peptide) - ion_position + 1)

  # Debugging print statements
  print(paste("ABC Ions DF Rows:", nrow(a_b_c_ions_df)))
  print(paste("YZ Ions DF Rows:", nrow(y_z_ions_df)))
  
  return(list(a_b_c_ions_df = a_b_c_ions_df, y_z_ions_df = y_z_ions_df))
}

# ions_result <- assemble_ions_df(matches_data_df_w_mods_updated, "983", "1")

check_unambiguity_before <- function(nearest_before, mod_position, a_b_c_ions_df, y_z_ions_df) {
  # If nearest_before is NA, return unambiguous
  if (is.na(nearest_before)) return(TRUE)

  abc_unambiguous <- any(a_b_c_ions_df$ion_position >= nearest_before & a_b_c_ions_df$ion_position < mod_position)
  yz_unambiguous <- any(y_z_ions_df$ion_position > nearest_before & y_z_ions_df$ion_position <= mod_position)
  return(abc_unambiguous || yz_unambiguous)
  # TRUE || NA will return TRUE, because if one operand is TRUE, the overall expression is TRUE regardless of the other operand
  # TRUE || FALSE will return TRUE, because if one operand is TRUE, the overall expression is TRUE regardless of the other operand
  # NA || TRUE will also return TRUE for the same reason
  # FALSE || NA will return NA, because the outcome depends on the unknown (NA) operand
  # NA || FALSE will return NA as well, for the same reason as above
  # NA || NA will return NA, because there's no sufficient information to determine the result
}

check_unambiguity_after <- function(nearest_after, mod_position, a_b_c_ions_df, y_z_ions_df) {
  # If nearest_after is NA, return unambiguous
  if (is.na(nearest_after)) return(TRUE)

  abc_unambiguous <- any(a_b_c_ions_df$ion_position < nearest_after & a_b_c_ions_df$ion_position >= mod_position)
  yz_unambiguous <- any(y_z_ions_df$ion_position <= nearest_after & y_z_ions_df$ion_position > mod_position)
  return(abc_unambiguous || yz_unambiguous)
}

# Add the required columns for ambiguity checks
matches_data_df_w_mods_updated <- matches_data_df_w_mods_updated %>%
  mutate(across(paste0("mod_pos_", 1:max_mods), 
                list(abc_before_unambiguous = ~NA,
                     yz_before_unambiguous = ~NA,
                     abc_after_unambiguous = ~NA,
                     yz_after_unambiguous = ~NA),
                .names = "{.col}_{.fn}"))

# Working function to apply ambiguity checks for each modification in a row
  apply_ambiguity_checks <- function(df, max_mods) {
  # Iterate through each mod_pos_ column
  for (i in 1:max_mods) {
    mod_pos_col <- paste0("mod_pos_", i)
    nearest_before_col <- paste0("nearest_before_mod_", i)
    nearest_after_col <- paste0("nearest_after_mod_", i)
    abc_before_unambiguous_col <- paste0("mod_pos_", i, "_abc_before_unambiguous")
    yz_before_unambiguous_col <- paste0("mod_pos_", i, "_yz_before_unambiguous")
    abc_after_unambiguous_col <- paste0("mod_pos_", i, "_abc_after_unambiguous")
    yz_after_unambiguous_col <- paste0("mod_pos_", i, "_yz_after_unambiguous")
    
    # Dynamically creating column names for assignment
    df <- df %>%
      rowwise() %>%
      mutate(
        !!abc_before_unambiguous_col := if_else(
          !is.na(.data[[mod_pos_col]]),
          # Assuming assemble_ions_df and check_unambiguity_before return a single value or NA_real_
          {
            ions <- assemble_ions_df(cur_data(), scan_number, PID)
            check_unambiguity_before(.data[[nearest_before_col]], .data[[mod_pos_col]], ions$a_b_c_ions_df, ions$y_z_ions_df)
          },
          NA_real_
        ),
        !!yz_before_unambiguous_col := if_else(
          !is.na(.data[[mod_pos_col]]),
          {
            ions <- assemble_ions_df(cur_data(), scan_number, PID)
            check_unambiguity_before(.data[[nearest_before_col]], .data[[mod_pos_col]], ions$a_b_c_ions_df, ions$y_z_ions_df)
          },
          NA_real_
        ),
        !!abc_after_unambiguous_col := if_else(
          !is.na(.data[[mod_pos_col]]),
          {
            ions <- assemble_ions_df(cur_data(), scan_number, PID)
            check_unambiguity_after(.data[[nearest_after_col]], .data[[mod_pos_col]], ions$a_b_c_ions_df, ions$y_z_ions_df)
          },
          NA_real_
        ),
        !!yz_after_unambiguous_col := if_else(
          !is.na(.data[[mod_pos_col]]),
          {
            ions <- assemble_ions_df(cur_data(), scan_number, PID)
            check_unambiguity_after(.data[[nearest_after_col]], .data[[mod_pos_col]], ions$a_b_c_ions_df, ions$y_z_ions_df)
          },
          NA_real_
        )
      ) %>%
      ungroup()
  }
  return(df)
}


# Make sure to pass the actual number of max_mods you have in your data
# test_df <- apply_ambiguity_checks(df, max_mods)

# Apply the checks
# matches_data_df_w_mods_updated <- apply_ambiguity_checks(matches_data_df_w_mods_updated)

# write_excel_csv(matches_data_df_w_mods_updated, "matches_data_df_w_mods_updated_ambiguity_check1")

```

Parallelization

```{r}
library(parallel)
library(dplyr)
library(tidyr)
library(stringr)

detectCores()

# Step 1: Group your dataframe by scan_number
grouped_df_list <- matches_data_df_w_mods_updated %>% 
  group_by(scan_number) %>%
  group_split()

# Step 2: Prepare to use parallel processing
numCores <- detectCores() / 2  # Reserve two cores for system tasks
cl <- makeCluster(numCores)
on.exit(stopCluster(cl), add = TRUE)

# Step 3: Export necessary objects and functions to each cluster node
clusterExport(cl, varlist = c("apply_ambiguity_checks", "max_mods", "assemble_ions_df", "check_unambiguity_before", "check_unambiguity_after"))
clusterEvalQ(cl, {
  library(dplyr)  # Assuming dplyr is used in your functions
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Execute the task in parallel
results <- parLapply(cl, grouped_df_list, function(group) {
  apply_ambiguity_checks(group, max_mods)
})

# Combine the results from all cores
# The method to combine results depends on the structure of your output
combined_results <- bind_rows(results)

combined_results <- combined_results %>%
  mutate(ion_type_position_charge = paste(ion_type, ion_position, ion_charge, sep = "_"))
```

Confirm whether a hit has disambiguating ions between all modification sites

```{r}
# Simplified check_columns function
disambiguate <- function(data, col_range_start, col_range_end) {
  # Assuming data is already filtered for a specific scan_number and PID
  
  # Dynamically select the range of columns if col_range_start and col_range_end are column names
  columns_to_check <- select(data, dplyr::all_of(col_range_start):dplyr::all_of(col_range_end))
  
  # Apply the check across the specified columns
  columns_check_result <- purrr::map_lgl(columns_to_check, function(column) {
    # Check if the column contains at least one "1"
    has_one <- any(column == 1, na.rm = TRUE)
    # If the column has at least one "1", it satisfies the condition
    if (has_one) {
      return(TRUE)
    } else {
      # If there are no "1"s, check if there's at least one NA (which also satisfies the condition)
      return(all(is.na(column)))
    }
  })
  
  # Return TRUE if all columns meet the criteria
  return(all(columns_check_result))
}

# Function to iterate over all scan_number and PID combinations and apply check_columns
disambiguate_all <- function(data, col_range_start, col_range_end) {
  results <- data %>%
    group_by(scan_number, PID) %>%
    summarise(all_true = disambiguate(cur_data(), col_range_start, col_range_end), .groups = 'drop')
  
  return(results)
}

# Step 1: Group your dataframe by scan_number
grouped_df_list <- combined_results %>% 
  group_by(scan_number) %>%
  group_split()

# Step 2: Prepare to use parallel processing
numCores <- detectCores() / 2  # Reserve two cores for system tasks
cl <- makeCluster(numCores)
on.exit(stopCluster(cl), add = TRUE)

# Step 3: Export necessary objects and functions to each cluster node
clusterExport(cl, varlist = c("disambiguate", "disambiguate_all"))
clusterEvalQ(cl, {
  library(dplyr)  # Assuming dplyr is used in your functions
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Execute the task in parallel
results <- parLapply(cl, grouped_df_list, function(group) {
  disambiguate_all(group, 'mod_pos_1_abc_before_unambiguous', 'mod_pos_7_yz_after_unambiguous')
})

# Combine the results from all cores
# The method to combine results depends on the structure of your output
disambiguated_results <- bind_rows(results)

# Assuming df1 is your original dataframe from which to extract peptide information
peptide_info_df <- combined_results %>%
  select(scan_number, PID, peptide) %>%
  distinct()  # Assuming you want unique combinations of scan_number, PID, and peptide

# Assuming df2 is the second dataframe to which you want to add the peptide information
disambiguated_results <- disambiguated_results %>%
  left_join(peptide_info_df, by = c("scan_number", "PID"))

disambiguated_true <- disambiguated_results %>%
  filter(all_true == TRUE)

disambiguated_filenames <- paste0()
```

Functions for extracting ion information from combined_results

```{r}
# Function to extract SHARED intensities from a given sublist
extract_intensities_from_sublist <- function(sublist, ions_list) {
  # Match ion_type_position with ions_list and extract intensity
  matched_intensities <- sublist %>%
    filter(ion_type_position_charge %in% ions_list) %>%
    pull(intensity)
  return(matched_intensities)
}

# Function to find the correct sublist based on scan_number and PID and extract intensities
extract_intensities <- function(scan_number, PID, ions_list, grouped_list) {
  # Find the sublist that matches scan_number and PID
  sublist <- grouped_list %>%
    purrr::keep(~ all(c(.x$scan_number == scan_number, .x$PID == PID))) %>%
    # Assuming there is only one match, otherwise you would need to handle multiple matches
    purrr::pluck(1)
  
  # If a matching sublist is found, extract intensities, else return numeric(0)
  if (!is.null(sublist) && nrow(sublist) > 0) {
    return(extract_intensities_from_sublist(sublist, ions_list))
  } else {
    return(numeric(0))
  }
}

# Function to extract shared ions from a given sublist
extract_ion_identities_from_sublist <- function(sublist, ions_list) {
  # Match ion_type_position with ions_list and extract intensity
  matched_identities <- sublist %>%
    filter(ion_type_position_charge %in% ions_list) %>%
    pull(ion_type_position_charge)
  return(matched_identities)
}

# Function to find the correct sublist based on scan_number and PID and extract intensities
extract_ordered_ion_lists <- function(scan_number, PID, ions_list, grouped_list) {
  # Find the sublist that matches scan_number and PID
  sublist <- grouped_list %>%
    purrr::keep(~ all(c(.x$scan_number == scan_number, .x$PID == PID))) %>%
    # Assuming there is only one match, otherwise you would need to handle multiple matches
    purrr::pluck(1)
  
  # If a matching sublist is found, extract intensities, else return numeric(0)
  if (!is.null(sublist) && nrow(sublist) > 0) {
    return(extract_ion_identities_from_sublist(sublist, ions_list))
  } else {
    return(numeric(0))
  }
}
```

Re-capturing ion lists for disambiguated hits

```{r}
# Assuming combined_results is your original dataframe from which to extract ion information
ion_info_df <- combined_results %>%
  group_by(scan_number, PID) %>%
  summarise(
    a_ions = list(unique(unlist(str_extract_all(ion_type_position_charge, "a(_nl_N_|_nl_O_|_nl_Kme3_|_nl_pST_|_nl_Ksu_|_)\\w+")))),
    b_ions = list(unique(unlist(str_extract_all(ion_type_position_charge, "b(_nl_N_|_nl_O_|_nl_Kme3_|_nl_pST_|_nl_Ksu_|_)\\w+")))),
    c_ions = list(unique(unlist(str_extract_all(ion_type_position_charge, "c(_nl_N_|_nl_O_|_nl_Kme3_|_nl_pST_|_nl_Ksu_|_)\\w+")))),
    y_ions = list(unique(unlist(str_extract_all(ion_type_position_charge, "y(_nl_N_|_nl_O_|_nl_Kme3_|_nl_pST_|_nl_Ksu_|_)\\w+")))),
    z_ions = list(unique(unlist(str_extract_all(ion_type_position_charge, "z(_nl_N_|_nl_O_|_nl_Kme3_|_nl_pST_|_nl_Ksu_|_)\\w+")))),
    .groups = 'drop'
  )

# Ensure unique(unlist()) is used to flatten the list from str_extract_all and remove duplicate matches before pasting

PSM_hits <- disambiguated_results %>%
  left_join(ion_info_df, by = c("scan_number", "PID"))

# set working directory
setwd(directory)

# Create dataframe of PEP values for merging to PSM_hits
PEP_df <- Byonic_df %>%
  select(Scan.Number.s...Posit., PID, PEP.2D)

names(PEP_df)[names(PEP_df) == "Scan.Number.s...Posit."] <- "scan_number"

# Add PEP values to PSM_hits
PSM_hits <- PSM_hits %>%
  left_join(PEP_df, by = c("scan_number", "PID"))

# Arrange by scan_number, then posterior_error_prob, and rank_order for tie-breaking
PSM_hits <- PSM_hits %>%
  group_by(scan_number) %>%
  arrange(scan_number, PEP.2D, PID, .by_group = TRUE) %>%
  ungroup()

# If you want to add a custom rank order based on this arrangement
PSM_hits <- PSM_hits %>%
  group_by(scan_number) %>%
  mutate(rank_order = row_number()) %>%
  ungroup()
```

Comparing ion lists

```{r}
# Create a dataframe with just the relevant ion information
scan_PID_ion_intensity <- combined_results %>%
  select("scan_number", "PID", "ion_type_position_charge", "intensity")

# Group by scan_number and PID for later searching
grouped_ion_intensity_list <- scan_PID_ion_intensity %>% 
  group_by(scan_number, PID) %>%
  group_split()

# Identify shared and unique ions for each PID
PSM_hits <- PSM_hits %>%
  # Group by scan_number and nest
  group_by(scan_number) %>%
  nest() %>%
  # Calculate shared ions and keep a combined list of all shared ions
  mutate(
    shared_ions_details = map(data, ~{
      a_ions_shared = Reduce(intersect, .x$a_ions)
      b_ions_shared = Reduce(intersect, .x$b_ions)
      c_ions_shared = Reduce(intersect, .x$c_ions)
      y_ions_shared = Reduce(intersect, .x$y_ions)
      z_ions_shared = Reduce(intersect, .x$z_ions)

      list(
        a_ions_shared = a_ions_shared,
        b_ions_shared = b_ions_shared,
        c_ions_shared = c_ions_shared,
        y_ions_shared = y_ions_shared,
        z_ions_shared = z_ions_shared,
        shared_ions = c(a_ions_shared, b_ions_shared, c_ions_shared, y_ions_shared, z_ions_shared) # Combined list of all shared ions
      )
    })
  ) %>%
  # Unnest the detailed shared ions for easier manipulation
  unnest_wider(shared_ions_details) %>%
  # Ensure we don't drop the nested data yet, as we'll need it for the join
  select(-data) %>%
  right_join(PSM_hits, by = "scan_number") %>%
  # Subtract shared ions to find unique ions for each type
  mutate(
    unique_a_ions = map2(a_ions, a_ions_shared, setdiff),
    unique_b_ions = map2(b_ions, b_ions_shared, setdiff),
    unique_c_ions = map2(c_ions, c_ions_shared, setdiff),
    unique_y_ions = map2(y_ions, y_ions_shared, setdiff),
    unique_z_ions = map2(z_ions, z_ions_shared, setdiff)
  )

PSM_hits <- PSM_hits %>%
  rowwise() %>%
  mutate(
        # Ensure non-NULL inputs and handle possibility of empty vectors
    unique_ions = list(c(unique_a_ions, unique_b_ions, unique_c_ions, unique_y_ions, unique_z_ions))
  ) %>%
  ungroup()

column_order = c("scan_number", "rank_order", "PEP.2D", "PID", "all_true", "peptide", "a_ions", "b_ions", "c_ions", "y_ions", "z_ions", "a_ions_shared", "b_ions_shared", "c_ions_shared", "y_ions_shared", "z_ions_shared", "shared_ions", "unique_a_ions", "unique_b_ions", "unique_c_ions", "unique_y_ions", "unique_z_ions", "unique_ions")
PSM_hits <- PSM_hits[, column_order]

# split hits for parallel processing: shared ion intensities
PSM_hits_list <- split(PSM_hits, PSM_hits$scan_number)
  
  # Initiate cluster
  no_cores <- detectCores() / 2  # Reserve two cores
  cl <- makeCluster(no_cores)
  on.exit(stopCluster(cl), add = TRUE)
  
  # Export necessary objects and functions
  clusterExport(cl, varlist = c("extract_intensities_from_sublist", "extract_intensities", "grouped_ion_intensity_list"))
  clusterEvalQ(cl, {
    library(dplyr)  # Assuming dplyr is used in your functions
    library(purrr)
    library(tidyr)
    library(stringr)
  })
  
  # Process the list of dataframes in parallel
  processed_chunks <- parLapply(cl, PSM_hits_list, function(df_chunk) {
    df_chunk %>%
      rowwise() %>%
      mutate(matched_intensities = list(extract_intensities(scan_number, PID, shared_ions, grouped_ion_intensity_list))) %>%
      ungroup()
  })
  
  # Stop the cluster
  stopCluster(cl)
  
  # Combine the processed chunks back into a single dataframe
  PSM_hits <- do.call(rbind, processed_chunks)

# split hits for parallel processing: shared ion intensities
PSM_hits_list <- split(PSM_hits, PSM_hits$scan_number)
  
  # Initiate cluster
  no_cores <- detectCores() / 2  # Reserve two cores
  cl <- makeCluster(no_cores)
  on.exit(stopCluster(cl), add = TRUE)

  # Export necessary objects and functions
  clusterExport(cl, varlist = c("extract_ion_identities_from_sublist", "extract_ordered_ion_lists", "grouped_ion_intensity_list"))
  clusterEvalQ(cl, {
    library(dplyr)  # Assuming dplyr is used in your functions
    library(purrr)
    library(tidyr)
    library(stringr)
  })
  
  # Process the list of dataframes in parallel
  processed_chunks <- parLapply(cl, PSM_hits_list, function(df_chunk) {
    df_chunk %>%
      rowwise() %>%
      mutate(ordered_ion_list = list(extract_ordered_ion_lists(scan_number, PID, shared_ions, grouped_ion_intensity_list))) %>%
      ungroup()
  })

  # Stop the cluster
  stopCluster(cl)
  
  # Combine the processed chunks back into a single dataframe
  PSM_hits <- do.call(rbind, processed_chunks)

# split hits for parallel processing: shared ion intensities
PSM_hits_list <- split(PSM_hits, PSM_hits$scan_number)
  
  # Initiate cluster
  no_cores <- detectCores() / 2  # Reserve two cores
  cl <- makeCluster(no_cores)
  on.exit(stopCluster(cl), add = TRUE)
  
  # Export necessary objects and functions
  clusterExport(cl, varlist = c("extract_intensities_from_sublist", "extract_intensities", "grouped_ion_intensity_list"))
  clusterEvalQ(cl, {
    library(dplyr)  # Assuming dplyr is used in your functions
    library(purrr)
    library(tidyr)
    library(stringr)
  })
  
  # Process the list of dataframes in parallel
  processed_chunks <- parLapply(cl, PSM_hits_list, function(df_chunk) {
    df_chunk %>%
      rowwise() %>%
      mutate(unique_ion_intensities = list(extract_intensities(scan_number, PID, unique_ions, grouped_ion_intensity_list))) %>%
      ungroup()
  })
  
  # Stop the cluster
  stopCluster(cl)
  
  # Combine the processed chunks back into a single dataframe
  PSM_hits <- do.call(rbind, processed_chunks)

# split hits for parallel processing: shared ion intensities
PSM_hits_list <- split(PSM_hits, PSM_hits$scan_number)
  
  # Initiate cluster
  no_cores <- detectCores() / 2  # Reserve two cores
  cl <- makeCluster(no_cores)
  on.exit(stopCluster(cl), add = TRUE)

  # Export necessary objects and functions
  clusterExport(cl, varlist = c("extract_ion_identities_from_sublist", "extract_ordered_ion_lists", "grouped_ion_intensity_list"))
  clusterEvalQ(cl, {
    library(dplyr)  # Assuming dplyr is used in your functions
    library(purrr)
    library(tidyr)
    library(stringr)
  })
  
  # Process the list of dataframes in parallel
  processed_chunks <- parLapply(cl, PSM_hits_list, function(df_chunk) {
    df_chunk %>%
      rowwise() %>%
      mutate(ordered_unique_ion_list = list(extract_ordered_ion_lists(scan_number, PID, unique_ions, grouped_ion_intensity_list))) %>%
      ungroup()
  })

  # Stop the cluster
  stopCluster(cl)
  
  # Combine the processed chunks back into a single dataframe
  PSM_hits <- do.call(rbind, processed_chunks)
```

Dividing PSM_hits by number of hits per scan and coefficient of variation

```{r}
# Add a column with the first rank_order's intensities for reference
PSM_hits <- PSM_hits %>%
  group_by(scan_number) %>%
  mutate(first_hit_intensities = list(matched_intensities[which.min(rank_order)])) %>%
  ungroup()

# Calculate ratios
PSM_hits <- PSM_hits %>%
  rowwise() %>%
  mutate(
    list(
      intensity_ratios = map2_dbl(
        matched_intensities, 
        first(first_hit_intensities), 
        ~ .x / .y
      )
    ) 
  )  %>%
  ungroup()

names(PSM_hits)[names(PSM_hits) == "list(...)"] <- "ratio"

# Step 2: Group by scan_number and rank_order, combine ratio values, and compute the average
PSM_hits <- PSM_hits %>%
  group_by(scan_number, rank_order) %>%
  mutate(
    combined_ratios = list(ratio),  # Combine ratios into a list
    average_ratio = mean(unlist(combined_ratios)[unlist(combined_ratios) != 1])  # Compute average of non-one values
  ) %>%
  ungroup()

# find coefficient of variance in average ratio
PSM_hits <- PSM_hits %>%
  group_by(scan_number, rank_order) %>%
  mutate(
    coefficient_of_variance = (sd(unlist(combined_ratios)[unlist(combined_ratios) != 1])/mean(unlist(combined_ratios)[unlist(combined_ratios) != 1]))  # Compute coefficient of variance of non-one values
  ) %>%
  ungroup()

file_name <- paste0("PSM_hits_", save_name_RData_suffix)
save(PSM_hits, file = file_name)

# subseting spectra based on number of unambiguous PSMs
# resolved when one assignment has a coefficient of variation that 
ten_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 10) %>%               # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

nine_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 9) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

eight_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 8) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

seven_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 7) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

six_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 6) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

five_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 5) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

four_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 4) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

three_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 3) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

two_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 2) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()

single_hit_scans <- PSM_hits %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 1) %>%                # Filter groups with more than one occurrence
  mutate(scan_hit = paste(scan_number, rank_order, sep = "_")) %>%                 
  ungroup()
```

Helper functions for processing scans with multiple hits

```{r}
# Helper functions

# Join hit dataframes
# When a hit dataframe is missing one ion type or match type in row one the column class gets lost
# This checks whether the class is character, and corrects it if not
conditional_merge_hit_df <- function(appended_df, main_df) {
  if (nrow(appended_df) > 0) {
    # Check and convert column types for appended_df
    if (class(appended_df$a_ions_shared) != "character") {
      appended_df$a_ions_shared <- as.character(appended_df$a_ions_shared)
    }
    if (class(appended_df$b_ions_shared) != "character") {
      appended_df$b_ions_shared <- as.character(appended_df$b_ions_shared)
    }
    if (class(appended_df$y_ions_shared) != "character") {
      appended_df$y_ions_shared <- as.character(appended_df$y_ions_shared)
    }
    if (class(appended_df$c_ions_shared) != "character") {
      appended_df$c_ions_shared <- as.character(appended_df$c_ions_shared)
    }
    if (class(appended_df$z_ions_shared) != "character") {
      appended_df$z_ions_shared <- as.character(appended_df$z_ions_shared)
    }
    
    # Check and convert column types for main_df
    if (class(main_df$a_ions_shared) != "character") {
      main_df$a_ions_shared <- as.character(main_df$a_ions_shared)
    }
    if (class(main_df$b_ions_shared) != "character") {
      main_df$b_ions_shared <- as.character(main_df$b_ions_shared)
    }
    if (class(main_df$y_ions_shared) != "character") {
      main_df$y_ions_shared <- as.character(main_df$y_ions_shared)
    }
    if (class(main_df$c_ions_shared) != "character") {
      main_df$c_ions_shared <- as.character(main_df$c_ions_shared)
    }
    if (class(main_df$z_ions_shared) != "character") {
      main_df$z_ions_shared <- as.character(main_df$z_ions_shared)
    }
    
    # Merge the data frames
    main_df <- main_df %>%
      bind_rows(appended_df)
  }
}
```

Processing three hit scans to attribute fractional intensity

```{r}
# First, create a nested dataframe grouped by scan_number
nested_df <- three_hit_scans %>%
  group_by(scan_number) %>%
  nest()

# Calculate the pairwise shared ions for each group and convert to comma-separated strings
nested_df <- nested_df %>%
  mutate(
    rank_12_shared_ions = map(data, ~ {
      ions_rank_1 <- .x$unique_ions[.x$rank_order == 1][[1]]
      ions_rank_2 <- .x$unique_ions[.x$rank_order == 2][[1]]
      intersect(ions_rank_1, ions_rank_2)
    }),
    rank_13_shared_ions = map(data, ~ {
      ions_rank_1 <- .x$unique_ions[.x$rank_order == 1][[1]]
      ions_rank_3 <- .x$unique_ions[.x$rank_order == 3][[1]]
      intersect(ions_rank_1, ions_rank_3)
    })
  )

# subsetting dataframe for binary comparisons between rank_orders
three_hit_one_two_scans <- three_hit_scans %>%
  filter(rank_order %in% c(1, 2)) %>%
  left_join(nested_df %>% select(scan_number, rank_12_shared_ions), by = "scan_number")
three_hit_one_three_scans <- three_hit_scans %>%
  filter(rank_order %in% c(1, 3)) %>%
  left_join(nested_df %>% select(scan_number, rank_13_shared_ions), by = "scan_number")

# split hits for parallel processing: unique ion intesities
three_hit_one_two_list <- split(three_hit_one_two_scans, three_hit_one_two_scans$scan_number)

# Initiate cluster
no_cores <- detectCores() / 2  # Reserve cores
cl <- makeCluster(no_cores)
on.exit(stopCluster(cl), add = TRUE)

# Export necessary objects and functions
clusterExport(cl, varlist = c("extract_intensities_from_sublist", "extract_intensities", "grouped_ion_intensity_list"))
clusterEvalQ(cl, {
  library(dplyr)  # Assuming dplyr is used in your functions
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Process the list of dataframes in parallel
processed_chunks <- parLapply(cl, three_hit_one_two_list, function(df_chunk) {
  df_chunk %>%
    rowwise() %>%
    mutate(rank_12_intensities = list(extract_intensities(scan_number, rank_order, rank_12_shared_ions, grouped_ion_intensity_list))) %>%
    ungroup()
})

# Stop the cluster
stopCluster(cl)

# Combine the processed chunks back into a single dataframe
three_hit_one_two_scans <- do.call(rbind, processed_chunks)

# split hits for parallel processing: unique ion intesities
three_hit_one_three_list <- split(three_hit_one_three_scans, three_hit_one_three_scans$scan_number)

# Initiate cluster
no_cores <- detectCores() / 2  # Reserve cores
cl <- makeCluster(no_cores)
on.exit(stopCluster(cl), add = TRUE)

# Export necessary objects and functions
clusterExport(cl, varlist = c("extract_intensities_from_sublist", "extract_intensities", "grouped_ion_intensity_list"))
clusterEvalQ(cl, {
  library(dplyr)  # Assuming dplyr is used in your functions
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Process the list of dataframes in parallel
processed_chunks <- parLapply(cl, three_hit_one_three_list, function(df_chunk) {
  df_chunk %>%
    rowwise() %>%
    mutate(rank_13_intensities = list(extract_intensities(scan_number, rank_order, rank_13_shared_ions, grouped_ion_intensity_list))) %>%
    ungroup()
})

# Stop the cluster
stopCluster(cl)

# Combine the processed chunks back into a single dataframe
three_hit_one_three_scans <- do.call(rbind, processed_chunks)

# Function to calculate ratios
three_hit_one_two_scans <- three_hit_one_two_scans %>%
  group_by(scan_number) %>%
  mutate(first_hit_intensities = list(rank_12_intensities[which.min(rank_order)])) %>%
  ungroup()

# Step 2: Calculate ratios row-wise
three_hit_one_two_scans <- three_hit_one_two_scans %>%
  rowwise() %>%
  mutate(
    intensity_ratios = list(
      map2(
        rank_12_intensities, 
        first(first_hit_intensities), 
        ~ .x / .y
      )
    )
  ) %>%
  ungroup()

# Function to calculate ratios
three_hit_one_three_scans <- three_hit_one_three_scans %>%
  group_by(scan_number) %>%
  mutate(first_hit_intensities = list(rank_13_intensities[which.min(rank_order)])) %>%
  ungroup()

# Step 2: Calculate ratios row-wise
three_hit_one_three_scans <- three_hit_one_three_scans %>%
  rowwise() %>%
  mutate(
    intensity_ratios = list(
      map2(
        rank_13_intensities, 
        first(first_hit_intensities), 
        ~ .x / .y
      )
    )
  ) %>%
  ungroup()

three_hit_one_two_scans <- three_hit_one_two_scans %>%
  rowwise() %>%  # Ensure operations are performed row by row
  mutate(
    average_12binary_ratio = {
      # Flatten the list of numeric vectors from intensity_ratios to a single numeric vector
      all_ratios <- unlist(intensity_ratios)
      # Add the value(s) from the 'ratio' column to all_ratios
      all_ratios_combined <- c(all_ratios, ratio)
      # Filter out ones from the combined ratios
      valid_ratios <- all_ratios_combined[all_ratios_combined != 1]
      # Calculate the average, if there are valid ratios available
      if (length(valid_ratios) > 0) {
        mean(valid_ratios)
      } else {
        NA_real_  # Return NA if no valid ratios
      }
    }
  ) %>%
  ungroup()  # Remove rowwise grouping

three_hit_one_three_scans <- three_hit_one_three_scans %>%
  rowwise() %>%  # Ensure operations are performed row by row
  mutate(
    average_13binary_ratio = {
      # Flatten the list of numeric vectors from intensity_ratios to a single numeric vector
      all_ratios <- unlist(intensity_ratios)
      # Add the value(s) from the 'ratio' column to all_ratios
      all_ratios_combined <- c(all_ratios, ratio)
      # Filter out ones from the combined ratios
      valid_ratios <- all_ratios_combined[all_ratios_combined != 1]
      # Calculate the average, if there are valid ratios available
      if (length(valid_ratios) > 0) {
        mean(valid_ratios)
      } else {
        NA_real_  # Return NA if no valid ratios
      }
    }
  ) %>%
  ungroup()  # Remove rowwise grouping

###
# Step 1:  calculate total intensity by row, omitting ions used to calculate the ratio between chimeric PSMs
# This 'total_intensity' is the sum of all non-unique ion intensities shared between multiple unambiguous PSMs
three_hit_one_two_scans <- three_hit_one_two_scans %>%
  rowwise() %>%
  mutate(total_shared_12_intensity = sum(rank_12_intensities[intensity_ratios == 1])) %>%
  ungroup()

# Step 2 & 3:  Group by 'scan_number' and calculate the minimum 'total_intensity' for each 'scan_number'
# The minimum total_shared_intensity 
min12_intensities <- three_hit_one_two_scans %>%
  group_by(scan_number) %>%
  summarise(min12_intensity = min(total_shared_12_intensity)) 

# Step 4: Join this minimum back to the original dataset
three_hit_one_two_scans <- three_hit_one_two_scans %>%
  left_join(min12_intensities, by = "scan_number")

# Filter for rank_order 2 and identify positions of non-one values in 'ratio'
positions_df <- three_hit_one_two_scans %>%
  filter(rank_order == 2) %>%
  mutate(non_one_positions = map(intensity_ratios, ~ which(.x != 1))) %>%
  select(scan_number, non_one_positions)

# and calculate the sum of matched_intensities at those positions
three_hit_one_two_scans <- three_hit_one_two_scans %>%
  left_join(positions_df, by = "scan_number") %>%
  mutate(
    sum_shared12_divergent_intensities = map2_dbl(rank_12_intensities, non_one_positions, ~ sum(.x[.y], na.rm = TRUE))
  )

###
# Step 1:  calculate total intensity by row, omitting ions used to calculate the ratio between chimeric PSMs
# This 'total_intensity' is the sum of all non-unique ion intensities shared between multiple unambiguous PSMs
three_hit_one_three_scans <- three_hit_one_three_scans %>%
  rowwise() %>%
  mutate(total_shared_13_intensity = sum(rank_13_intensities[intensity_ratios == 1])) %>%
  ungroup()

# Step 2 & 3:  Group by 'scan_number' and calculate the minimum 'total_intensity' for each 'scan_number'
# The minimum total_shared_intensity 
min13_intensities <- three_hit_one_three_scans %>%
  group_by(scan_number) %>%
  summarise(min13_intensity = min(total_shared_13_intensity)) 

# Step 4: Join this minimum back to the original dataset
three_hit_one_three_scans <- three_hit_one_three_scans %>%
  left_join(min13_intensities, by = "scan_number")

# Filter for rank_order 2 and identify positions of non-one values in 'ratio'
positions_df <- three_hit_one_three_scans %>%
  filter(rank_order == 3) %>%
  mutate(non_one_positions = map(intensity_ratios, ~ which(.x != 1))) %>%
  select(scan_number, non_one_positions)

# and calculate the sum of matched_intensities at those positions
three_hit_one_three_scans <- three_hit_one_three_scans %>%
  left_join(positions_df, by = "scan_number") %>%
  mutate(
    sum_shared13_divergent_intensities = map2_dbl(rank_13_intensities, non_one_positions, ~ sum(.x[.y], na.rm = TRUE))
  )

# Allocating unattributed intensity
# for binary chimeric mixtures
# total signal = signal A + signal B
# B/A ratio = x
# total signal = (1 + x) signal A
# signal A = total signal / (1 + x)
# total signal = (1 + (1/x)) signal B
# signal B = total signal / (1 + (1/x))

three_hit_one_two_scans <- three_hit_one_two_scans %>%
  mutate(bin12_ratio = ifelse(is.na(average_12binary_ratio), 1, average_12binary_ratio))
# establishing ratio specifically for apportioning fractional intensity in this df

three_hit_one_two_scans <- three_hit_one_two_scans %>%
  group_by(scan_number) %>%
  mutate(
    fractional_12_intensity = case_when(
      rank_order == 1 ~ min12_intensity / (1 + bin12_ratio),
      rank_order == 2 ~ min12_intensity / (1 + (1 / bin12_ratio)),
      TRUE ~ NA_real_ # for rank_order beyond the first two, or adjust as needed
    )
  ) %>%
  ungroup() # Always good practice to ungroup after you're done

three_hit_one_three_scans <- three_hit_one_three_scans %>%
  mutate(bin13_ratio = ifelse(is.na(average_13binary_ratio), 1, average_13binary_ratio))
# establishing ratio specifically for apportioning fractional intensity in this df

three_hit_one_three_scans <- three_hit_one_three_scans %>%
  group_by(scan_number) %>%
  mutate(
    fractional_13_intensity = case_when(
      rank_order == 1 ~ min13_intensity / (1 + bin13_ratio),
      rank_order == 3 ~ min13_intensity / (1 + (1 / bin13_ratio)),
      TRUE ~ NA_real_ # for rank_order beyond the first two, or adjust as needed
    )
  ) %>%
  ungroup() # Always good practice to ungroup after you're done

```

Fractional signal intensity attribution for three hit spectra

```{r}
three_hit_scans <- three_hit_scans %>%
  left_join(three_hit_one_two_scans %>% select(scan_number, rank_order, rank_12_shared_ions, min12_intensity, sum_shared12_divergent_intensities, fractional_12_intensity, average_12binary_ratio), by = c("scan_number" = "scan_number", "rank_order" = "rank_order"))  %>%
  left_join(three_hit_one_three_scans %>% select(scan_number, rank_order, rank_13_shared_ions, min13_intensity, sum_shared13_divergent_intensities, fractional_13_intensity, average_13binary_ratio), by = c("scan_number" = "scan_number", "rank_order" = "rank_order"))

three_hit_scans <- three_hit_scans %>%
  rowwise() %>%
  mutate(
    combined_shared_ions = list(c(
      if (length(rank_12_shared_ions) > 0) rank_12_shared_ions else character(0),
      if (length(rank_13_shared_ions) > 0) rank_13_shared_ions else character(0)
    )),
    remaining_unique_ions = list(setdiff(unique_ions, unlist(combined_shared_ions)))
  ) %>%
  ungroup() %>%
  select(-combined_shared_ions) # Optional: Remove the intermediate combined_shared_ions column if not needed

# Adding ion intensities for remaining unique ions
# split hits for parallel processing: shared ion intensities
three_hit_scans_list <- split(three_hit_scans, three_hit_scans$scan_number)

# Initiate cluster
no_cores <- detectCores() / 2  # Reserve one core
cl <- makeCluster(no_cores)
on.exit(stopCluster(cl), add = TRUE)

# Export necessary objects and functions
clusterExport(cl, varlist = c("extract_intensities_from_sublist", "extract_intensities", "grouped_ion_intensity_list"))
clusterEvalQ(cl, {
  library(dplyr)  # Assuming dplyr is used in your functions
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Process the list of dataframes in parallel
processed_chunks <- parLapply(cl, three_hit_scans_list, function(df_chunk) {
  df_chunk %>%
    rowwise() %>%
    mutate(remaining_unique_ion_intensities = list(extract_intensities(scan_number, rank_order, remaining_unique_ions, grouped_ion_intensity_list))) %>%
    ungroup()
})

# Stop the cluster
stopCluster(cl)

# Combine the processed chunks back into a single dataframe
three_hit_scans <- do.call(rbind, processed_chunks)

# Define the custom function
retrieve_value <- function(rank_order, average_12binary_ratio, average_13binary_ratio, average_ratio) {
  # Initialize the value to retrieve
  value_to_retrieve <- NA
  
  # Direct use of coalesce depends on rank_order
  if (rank_order == 1) {
    value_to_retrieve <- coalesce(average_13binary_ratio, average_12binary_ratio, average_ratio, 1)
  } else if (rank_order == 2) {
    # Adjust the order or selection of columns as needed for rank_order == 2
    value_to_retrieve <- coalesce(average_12binary_ratio, average_ratio, 1)
  } else if (rank_order == 3) {
    # Adjust the order or selection of columns as needed for rank_order == 3
    value_to_retrieve <- coalesce(average_13binary_ratio, average_ratio, 1)
  } else {
    # Default case or error handling
    value_to_retrieve <- NA_real_
  }
  
  return(value_to_retrieve)
}

# Apply the function row-wise on your dataframe
three_hit_scans <- three_hit_scans %>%
  rowwise() %>%
  mutate(retrieved_value = retrieve_value(rank_order, average_12binary_ratio, average_13binary_ratio, average_ratio)) %>%
  ungroup()

# Calculating fractional intensity
# Filter for rank_order 2 and identify positions of non-one values in 'ratio'
positions_df <- three_hit_scans %>%
  filter(rank_order == 2) %>%
  mutate(non_one_pos_rank2 = map(ratio, ~ which(.x != 1))) %>%
  select(scan_number, non_one_pos_rank2)

# Join position values back to dataframe
three_hit_scans <- three_hit_scans %>%
  left_join(positions_df, by = "scan_number")

# Filter for rank_order 3 and identify positions of non-one values in 'ratio'
positions_df <- three_hit_scans %>%
  filter(rank_order == 3) %>%
  mutate(non_one_pos_rank3 = map(ratio, ~ which(.x != 1))) %>%
  select(scan_number, non_one_pos_rank3)

# Join position values back to dataframe
three_hit_scans <- three_hit_scans %>%
  left_join(positions_df, by = "scan_number")

# Merge position value columns
three_hit_scans <- three_hit_scans %>%
  mutate(
    non_one_positions = map2(non_one_pos_rank2, non_one_pos_rank3, ~{
      combined <- unique(c(.x, .y))
      if (length(combined) == 0) 0 else combined
    })
  )
# This shows us which ions were unique to one of the hits in the three unambiguous hits

# and calculate the sum of matched_intensities at those positions
three_hit_scans <- three_hit_scans %>%
  rowwise() %>%
  mutate(
    sum_non_one_ratio_intensities = sum(matched_intensities[unlist(non_one_positions)], na.rm = TRUE)
  ) %>%
  ungroup()

# We need to calculate total intensity that must be divided among the three unambiguous hits
three_hit_scans <- three_hit_scans %>%
  rowwise() %>%
  mutate(
    # Check for 0 and handle accordingly for matched_intensities
    sum_matched_excluded = if (all(non_one_positions == 0)) {
      sum(matched_intensities, na.rm = TRUE)  # If all values are 0, sum all intensities
    } else {
      sum(matched_intensities[-non_one_positions], na.rm = TRUE)
    },
    # Check for 0 and handle accordingly for first_hit_intensities
    sum_first_hit_excluded = if (all(non_one_positions == 0)) {
      sum(first_hit_intensities[[1]], na.rm = TRUE)  # If all values are 0, sum all intensities
    } else {
      sum(first_hit_intensities[[1]][-non_one_positions], na.rm = TRUE)
    }
  ) %>%
  ungroup()
# For those values that attributable to only two of the three we will have to do another calculation  

# Spread the ratios into separate columns for each rank_order
three_hit_scans_wide <- three_hit_scans %>%
  select(scan_number, rank_order, retrieved_value) %>%
  pivot_wider(
    names_from = rank_order,
    values_from = retrieved_value,
    names_prefix = "rank_order",
    values_fill = list(retrieved_value = NA),  # Fill missing values with NA
    # Apply a function to replace NaN with 1
    values_fn = list(retrieved_value = function(x) ifelse(is.nan(x), 1, x))
  )

# Join the wide format ratios back to the original dataframe
three_hit_scans <- three_hit_scans %>%
  left_join(three_hit_scans_wide, by = "scan_number")

# for ternary chimeric mixtures
# total signal = signal A + signal B + signal C
# B/A ratio = x; C/A ratio = y
# total signal = (1 + x + y) signal A
# total signal = (1 + x + y) * (signal B / x)
# total signal = ((1/x) + 1 + (x/y)) * signal B
# total signal = (1 + x + y) * (signal C / y)
# total signal = ((1/y) + (y/x) + 1) * signal C

# Calculating fractional intensity
three_hit_scans <- three_hit_scans %>%
  rowwise() %>%
  mutate(
    fractional_intensity = case_when(
      rank_order == 1 ~ sum_matched_excluded / sum(rank_order1, rank_order2, rank_order3),
      rank_order == 2 ~ sum_matched_excluded / sum(rank_order1, (1 / rank_order2), (rank_order3 / rank_order2)),
      rank_order == 3 ~ sum_matched_excluded / sum(rank_order1, (1 / rank_order3), (rank_order2 / rank_order3)),
      TRUE ~ NA_real_ # for rank_order beyond the first two, or adjust as needed
    )
  )

# Calculate summed intensities from fractional_intensity, sum_matched_intensities and sum of unique_ion_intensities 
three_hit_scans <- three_hit_scans %>%
  rowwise() %>%
  mutate(total_intensity = sum(
    sum_shared12_divergent_intensities,
    fractional_12_intensity,
    sum_shared13_divergent_intensities,
    fractional_13_intensity,
    sum(remaining_unique_ion_intensities),
    sum_non_one_ratio_intensities,
    fractional_intensity, na.rm = TRUE)
  ) %>%
  ungroup()
```

Fractional signal intensity attribution for two hit scans

```{r}
two_hit_scans <- two_hit_scans[order(two_hit_scans$scan_number, decreasing = FALSE),]

# remove coefficient of variance > 0.5
two_hit_scans <- two_hit_scans %>%
  filter(coefficient_of_variance < 0.5 | is.na(coefficient_of_variance))

# resolved when one assignment has a coefficient of variation that 
single_result <- two_hit_scans %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 1) %>%                 # Filter groups with more than one occurrence
  ungroup()             

  # Merge the data frames
  single_hit_scans <- conditional_merge_hit_df(single_hit_scans, single_result)

# resolved when one assignment has a coefficient of variation that 
two_hit_scans <- two_hit_scans %>%
  group_by(scan_number) %>%           # Group data by the column of interest
  filter(n() == 2) %>%                 # Filter groups with more than one occurrence
  ungroup()

# Step 1:  calculate total intensity by row, omitting ions used to calculate the ratio between chimeric PSMs
# This 'total_intensity' is the sum of all non-unique ion intensities shared between multiple unambiguous PSMs
two_hit_scans <- two_hit_scans %>%
  rowwise() %>%
  mutate(total_shared_intensity = sum(matched_intensities[ratio == 1])) %>%
  ungroup()

# Step 2 & 3:  Group by 'scan_number' and calculate the minimum 'total_intensity' for each 'scan_number'
# The minimum total_shared_intensity 
min_intensities <- two_hit_scans %>%
  group_by(scan_number) %>%
  summarise(min_intensity = min(total_shared_intensity)) 

# Step 4: Join this minimum back to the original dataset
two_hit_scans <- two_hit_scans %>%
  left_join(min_intensities, by = "scan_number")

# Calculate the non-one 'ratio' for each 'scan_number' and replace NaN with 1
# All the 1:1 ratios came through a prior calculation as NaN
# It might be better to correct this elsewhere, but leaving it in place up until now lets us pull out the averge ratio values with min even when those ratios are greater than 1
min_average_ratio <- two_hit_scans %>%
  group_by(scan_number) %>%
  summarise(min_average_ratio = min(average_ratio, na.rm = TRUE)) %>%
  mutate(min_average_ratio = if_else(is.infinite(min_average_ratio), 1, min_average_ratio))

# Join this minimum back to the original dataset
two_hit_scans <- two_hit_scans %>%
  left_join(min_average_ratio, by = "scan_number")

# assign rank_orders
two_hit_scans <- two_hit_scans %>%
  group_by(scan_number) %>%
  mutate(rank_order = row_number(rank_order)) %>%
  ungroup() # Always good practice to ungroup after you're done

# for binary chimeric mixtures
# total signal = signal A + signal B
# B/A ratio = x
# total signal = (1 + x) signal A
# signal A = total signal / (1 + x)
# total signal = (1 + (1/x)) signal B
# signal B = total signal / (1 + (1/x))

# Calculating fractional intensity
two_hit_scans <- two_hit_scans %>%
  group_by(scan_number) %>%
  mutate(
    fractional_intensity = case_when(
      rank_order == 1 ~ min_intensity / (1 + min_average_ratio),
      rank_order == 2 ~ min_intensity / (1 + (1 / min_average_ratio)),
      TRUE ~ NA_real_ # for rank_order beyond the first two, or adjust as needed
    )
  ) %>%
  ungroup() # Always good practice to ungroup after you're done

# Filter for rank_order 2 and identify positions of non-one values in 'ratio'
positions_df <- two_hit_scans %>%
  filter(rank_order == 2) %>%
  mutate(non_one_positions = map(ratio, ~ which(.x != 1))) %>%
  select(scan_number, non_one_positions)

# and calculate the sum of matched_intensities at those positions
two_hit_scans <- two_hit_scans %>%
  left_join(positions_df, by = "scan_number") %>%
  mutate(
    sum_matched_intensities = map2_dbl(matched_intensities, non_one_positions, ~ sum(.x[.y], na.rm = TRUE))
  )

# Calculate summed intensities from fractional_intensity, sum_matched_intensities and sum of unique_ion_intensities 
two_hit_scans <- two_hit_scans %>%
  rowwise() %>%
  mutate(total_intensity = sum(sum(unique_ion_intensities), fractional_intensity, sum_matched_intensities)) %>%
  ungroup()
```

Quantifying unambiguous hits

```{r}
# Order entries
single_hit_scans <- single_hit_scans[order(single_hit_scans$scan_number, decreasing = FALSE),]

# Adding matched ion intensities to unambiguous hits and calculating total ion intensity
single_hit_scans <- single_hit_scans %>%
  rowwise() %>%
  mutate(total_intensity = sum(matched_intensities)) %>%
  ungroup()
```

Merge hit dataframes

```{r}
column_order = c("scan_number", "rank_order", "PEP.2D", "PID", "all_true", "peptide", "a_ions", "b_ions", "c_ions", "y_ions", "z_ions", "all_ions", "matched_intensities", "total_intensity")

two_hit_scans <- two_hit_scans %>%
  rowwise() %>%
  mutate(
    # Ensure non-NULL inputs and handle possibility of empty vectors
    all_ions = list(c(a_ions, b_ions, c_ions, y_ions, z_ions))
  ) %>%
  mutate(matched_intensities = list(extract_intensities(scan_number, rank_order, all_ions, grouped_ion_intensity_list))) %>%
  ungroup()

two_hit_scans_backup <- two_hit_scans

two_hit_scans <- two_hit_scans[, column_order]

three_hit_scans_backup <- three_hit_scans

three_hit_scans <- three_hit_scans %>%
  rowwise() %>%
  mutate(
    # Ensure non-NULL inputs and handle possibility of empty vectors
    all_ions = list(c(a_ions, b_ions, c_ions, y_ions, z_ions))
  )

# split hits for parallel processing: shared ion intensities
three_hit_scans_list <- split(three_hit_scans, three_hit_scans$scan_number)

# Initiate cluster
no_cores <- detectCores() / 2  # Reserve one core
cl <- makeCluster(no_cores)
on.exit(stopCluster(cl), add = TRUE)

# Export necessary objects and functions
clusterExport(cl, varlist = c("extract_intensities_from_sublist", "extract_intensities", "grouped_ion_intensity_list"))
clusterEvalQ(cl, {
  library(dplyr)  # Assuming dplyr is used in your functions
  library(purrr)
  library(tidyr)
  library(stringr)
})

# Process the list of dataframes in parallel
processed_chunks <- parLapply(cl, three_hit_scans_list, function(df_chunk) {
  df_chunk %>%
    rowwise() %>%
    mutate(matched_intensities = list(extract_intensities(scan_number, rank_order, all_ions, grouped_ion_intensity_list))) %>%
    ungroup()
})

# Stop the cluster
stopCluster(cl)

# Combine the processed chunks back into a single dataframe
three_hit_scans <- do.call(rbind, processed_chunks)

three_hit_scans <- three_hit_scans[, column_order]

single_hit_scans <- single_hit_scans %>%
  rowwise() %>%
  mutate(
    # Ensure non-NULL inputs and handle possibility of empty vectors
    all_ions = list(c(a_ions, b_ions, c_ions, y_ions, z_ions))
  )

single_hit_scans <- single_hit_scans[, column_order]

assigned_fractional_intensity_hits <- single_hit_scans %>%
  bind_rows(two_hit_scans)

assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  bind_rows(three_hit_scans)

assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits[order(assigned_fractional_intensity_hits$scan_number, decreasing = FALSE),]

combined_results_column_order <- c("scan_number",
                                   "PID",
                                   "mod_pos_1",
                                   "mod_mass_1",
                                   "mod_pos_2",
                                   "mod_mass_2",
                                   "mod_pos_3",
                                   "mod_mass_3",
                                   "mod_pos_4",
                                   "mod_mass_4",
                                   "mod_pos_5",
                                   "mod_mass_5",
                                   "mod_pos_6",
                                   "mod_mass_6",
                                   "mod_pos_7",
                                   "mod_mass_7")

combined_mod_results <- combined_results[, combined_results_column_order]

combined_mod_results <- combined_mod_results %>%
  mutate(scan_PID = paste(scan_number, PID, sep = "_")) %>%
  distinct(scan_PID, .keep_all = TRUE)

assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  left_join(combined_mod_results, by = c("scan_number", "PID")) %>%
  rowwise() %>%
  ungroup()
```

Formatting by PTM type and location

```{r}
# Example lookup dataframe
lookup_PTM_df <- data.frame(
  num_value = c(14.0157, 28.0313, 42.0470, 42.0106, 56.0262, 79.9663, 86.0368, 100.0160, 72.0211, 305.2630), # Your numerical values
  text_string = c("me1", "me2", "me3", "ac", "prop", "phos", "hb", "su", "lac", "srt") # Corresponding text descriptions
)

site_df <- data.frame(
  num_value = c(2, 3, 4, 5, 6, 8, 9, 10, 11, 14, 17, 18, 19, 22, 23, 26, 27, 28, 31, 32, 36), # Your numerical values
  text_string = c("R", "T", "K", "Q", "T", "R", "K","S", "T", "K", "R", "K", "Q", "T", "K", "R", "K", "S", "S", "T", "Cterm") # Corresponding text descriptions
)

# Assuming the function checks a single value against the ref_df
check_against_reference <- function(input_value, ref_df) {
  if (is.na(input_value)) {
    return("")  # Or return("") for an empty cell
  } else {
    matched_entry <- ref_df %>%
      filter(num_value == input_value) %>%  # Ensure 'num_value' exactly matches the column name in 'ref_df'
      pull(text_string)  # Assuming 'text_string' is the column you want to retrieve
    
    if (length(matched_entry) == 0) {
      return("")
    } else {
      return(matched_entry)
    }
  }
}

file_name <- paste0("experiment1_all_hits_assigned_fractional_intensity_hits_", save_name_RData_suffix)
save(assigned_fractional_intensity_hits, file = file_name)

assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  rowwise() %>%
  mutate(
    resi1 = check_against_reference(mod_pos_1, site_df),
    resi2 = check_against_reference(mod_pos_2, site_df),
    resi3 = check_against_reference(mod_pos_3, site_df),
    resi4 = check_against_reference(mod_pos_4, site_df),
    resi5 = check_against_reference(mod_pos_5, site_df),
    resi6 = check_against_reference(mod_pos_6, site_df),
    resi7 = check_against_reference(mod_pos_7, site_df),
    mod1 = check_against_reference(mod_mass_1, lookup_PTM_df),
    mod2 = check_against_reference(mod_mass_2, lookup_PTM_df),
    mod3 = check_against_reference(mod_mass_3, lookup_PTM_df),
    mod4 = check_against_reference(mod_mass_4, lookup_PTM_df),
    mod5 = check_against_reference(mod_mass_5, lookup_PTM_df),
    mod6 = check_against_reference(mod_mass_6, lookup_PTM_df),
    mod7 = check_against_reference(mod_mass_7, lookup_PTM_df),
  ) %>%
  mutate(
    PTM1 = paste(resi1, mod_pos_1, mod1, sep = ""),
    PTM2 = paste(resi2, mod_pos_2, mod2, sep = ""),
    PTM3 = paste(resi3, mod_pos_3, mod3, sep = ""),
    PTM4 = paste(resi4, mod_pos_4, mod4, sep = ""),
    PTM5 = paste(resi5, mod_pos_5, mod5, sep = ""),
    PTM6 = paste(resi6, mod_pos_6, mod6, sep = ""),
    PTM7 = paste(resi7, mod_pos_7, mod7, sep = "")
  ) %>%
  mutate(across(c(PTM1, PTM2, PTM3, PTM4, PTM5, PTM6, PTM7), ~ifelse(. == "NA", NA, .))) %>%
  mutate(all_mods = pmap_chr(list(PTM1, PTM2, PTM3, PTM4, PTM5, PTM6, PTM7), function(...) {
    # Combine non-NA elements into a single string
    paste(na.omit(c(...)), collapse = " ")
  })) %>%
  ungroup()

unique_proteoforms <-  as.vector(unique(assigned_fractional_intensity_hits$all_mods, .keep_all = TRUE))

# flatten list columns to strings for preservation when exporting to excel
assigned_hits_export <- assigned_fractional_intensity_hits %>%
  mutate(across(where(is.list), ~sapply(., toString)))

setwd(directory)
file_name <- paste0("experiment1_all_hits_assigned_fractional_intensity_hits_", save_name_csv_suffix)
write_excel_csv(assigned_hits_export, file = file_name)

H31_assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  filter(str_detect(peptide, "ARTKQTARKSTGGKAPRKQLATKAARKSAPATGGGH"))

H33_assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  filter(str_detect(peptide, "ARTKQTARKSTGGKAPRKQLATKAARKSAPSTGGGH"))

# Another approach is to serialize list-containing rows as JSON strings. JSON can preserve the nested structure of lists and is more suited for conveying complex data structures than CSV
#library(jsonlite)

# Convert each row to a JSON string
#json_strings <- apply(assigned_fractional_intensity_hits, 1, toJSON, auto_unbox = TRUE)

# If you want to put these strings in a dataframe and then write to CSV for Excel:
#json_df <- data.frame(json = json_strings)
```

TMT calculations

```{r}
assigned_fractional_intensity_backup <- assigned_fractional_intensity_hits
assigned_fractional_intensity_hits <- assigned_fractional_intensity_backup

# Function for extracting TMT ion intensities
find_intensities_for_scan <- function(scan_number, mz, target_mz, ppm_tolerance) {
  # Convert the scan number to an integer if necessary
  scan_number_int <- as.integer(scan_number)
  
  # Retrieve the peaks data for the given scan number from the mz object
  scan_data <- peaks(mz, scan = scan_number_int)
  
  # Proceed to find intensities within the tolerance for the target m/z values
  intensities <- sapply(target_mz, function(mz) {
    tolerance <- calculate_tolerance(mz, ppm_tolerance)
    idx <- which(scan_data[, 1] >= mz - tolerance & scan_data[, 1] <= mz + tolerance)
    
    if (length(idx) > 0) {
      # Calculate distances for all matches
      distances <- abs(scan_data[idx, 1] - mz)
      # Select the index with the smallest distance
      closest_match_idx <- idx[which.min(distances)]
      return(scan_data[closest_match_idx, 2])  # Returning the intensity of the closest match
    } else {
      return(NA)  # Return NA if no matches are found
    }
  })
  
  return(intensities)
}

mz <- experiment1_mz

# Extract raw TMT signals
assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  rowwise() %>%
  mutate('126_intensity' = as.numeric(find_intensities_for_scan(scan_number, mz, TMT_ETD_mz_values[1], ppm_tolerance)),
         '127_intensity' = as.numeric(find_intensities_for_scan(scan_number, mz, TMT_ETD_mz_values[2], ppm_tolerance)),
         '128_intensity' = as.numeric(find_intensities_for_scan(scan_number, mz, TMT_ETD_mz_values[3], ppm_tolerance)),
         '129_intensity' = as.numeric(find_intensities_for_scan(scan_number, mz, TMT_ETD_mz_values[4], ppm_tolerance)),
         '130_intensity' = as.numeric(find_intensities_for_scan(scan_number, mz, TMT_ETD_mz_values[5], ppm_tolerance)),
         '131_intensity' = as.numeric(find_intensities_for_scan(scan_number, mz, TMT_ETD_mz_values[6], ppm_tolerance))
  )

# Function to identify rows with => 5 TMT signals
TMT_check <- function(data, col_range_start, col_range_end, TMT_count) {
  # Dynamically select the range of columns
  columns_to_check <- data %>% select(dplyr::all_of(col_range_start):dplyr::all_of(col_range_end))
  
  # Apply the check across the specified columns and count how many satisfy the condition
  columns_check_result <- purrr::map_lgl(columns_to_check, ~any(.x > 0, na.rm = TRUE))
  
  # Return TRUE if at least 5 columns meet the criteria
  sum(columns_check_result) >= TMT_count
}

# Identifying rows with => 5 TMT signals
assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  rowwise() %>%
  mutate('5_TMT' = TMT_check(cur_data(), '126_intensity', '131_intensity', 5)) %>%
  mutate('6_TMT' = TMT_check(cur_data(), '126_intensity', '131_intensity', 6)) %>%
  ungroup()

# Applying isotopic corrections from manufacturer CoA from TMT batch
assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  rowwise() %>%
  mutate(
    `126_corrected` = (`126_intensity` * TMT126[3]) - sum(c(`127_intensity` * TMT126[4],
                                                            `128_intensity` * TMT126[5]),
                                                          na.rm = TRUE),
    `127_corrected` = (`127_intensity` * TMT127[3]) - sum(c(`126_intensity` * TMT127[2],
                                                            `128_intensity` * TMT127[4],
                                                            `129_intensity` * TMT127[5]),
                                                          na.rm = TRUE),
    `128_corrected` = (`128_intensity` * TMT128[3]) - sum(c(`126_intensity` * TMT128[1],
                                                            `127_intensity` * TMT128[2],
                                                            `129_intensity` * TMT128[4],
                                                            `130_intensity` * TMT128[5]),
                                                          na.rm = TRUE),
    `129_corrected` = (`129_intensity` * TMT129[3]) - sum(c(`127_intensity` * TMT129[1],
                                                            `128_intensity` * TMT129[2],
                                                            `130_intensity` * TMT129[4],
                                                            `131_intensity` * TMT129[5]),
                                                          na.rm = TRUE),
    `130_corrected` = (`130_intensity` * TMT130[3]) - sum(c(`128_intensity` * TMT130[1],
                                                            `129_intensity` * TMT130[2],
                                                            `131_intensity` * TMT130[4]),
                                                          na.rm = TRUE),
    `131_corrected` = (`131_intensity` * TMT131[3]) - sum(c(`129_intensity` * TMT131[1],
                                                            `130_intensity` * TMT131[2]),
                                                          na.rm = TRUE)
  )

# Remove duplicate scans for calculating channel mean, median, etc.
TMT_single_scan_df <- assigned_fractional_intensity_hits %>%
  distinct(scan_number, .keep_all = TRUE)
```

TMT signal normalization

```{r}
# CUMULATIVE NORMALIZATION
# Calculating individual channel intensities for PSMs
PSM_TMT126_total_channel_signal <- sum(TMT_single_scan_df$`126_corrected`, na.rm = TRUE)
PSM_TMT127_total_channel_signal <- sum(TMT_single_scan_df$`127_corrected`, na.rm = TRUE)
PSM_TMT128_total_channel_signal <- sum(TMT_single_scan_df$`128_corrected`, na.rm = TRUE)
PSM_TMT129_total_channel_signal <- sum(TMT_single_scan_df$`129_corrected`, na.rm = TRUE)
PSM_TMT130_total_channel_signal <- sum(TMT_single_scan_df$`130_corrected`, na.rm = TRUE)
PSM_TMT131_total_channel_signal <- sum(TMT_single_scan_df$`131_corrected`, na.rm = TRUE)

# Calculating channel normalization factors
cumulative_channel_signals <- c(PSM_TMT126_total_channel_signal,
                                PSM_TMT127_total_channel_signal,
                                PSM_TMT128_total_channel_signal,
                                PSM_TMT129_total_channel_signal,
                                PSM_TMT130_total_channel_signal,
                                PSM_TMT131_total_channel_signal)

max_cumulative_signal <- max(cumulative_channel_signals)

cumulative_channel_norm <- (cumulative_channel_signals / max_cumulative_signal)

# Normalizing individual channel intensities for PSMs
TMT_single_scan_df <- TMT_single_scan_df %>%
  rowwise() %>%
  mutate( `126_cum_normalized` = `126_corrected` / cumulative_channel_norm[1],
          `127_cum_normalized` = `127_corrected` / cumulative_channel_norm[2],
          `128_cum_normalized` = `128_corrected` / cumulative_channel_norm[3],
          `129_cum_normalized` = `129_corrected` / cumulative_channel_norm[4],
          `130_cum_normalized` = `130_corrected` / cumulative_channel_norm[5],
          `131_cum_normalized` = `131_corrected` / cumulative_channel_norm[6],
  ) %>%
  ungroup()

# log2 transform
TMT_single_scan_df <- TMT_single_scan_df %>%
  rowwise() %>%
  mutate( log2_cum_126 = log2(`126_cum_normalized`),
          log2_cum_127 = log2(`127_cum_normalized`),
          log2_cum_128 = log2(`128_cum_normalized`),
          log2_cum_129 = log2(`129_cum_normalized`),
          log2_cum_130 = log2(`130_cum_normalized`),
          log2_cum_131 = log2(`131_cum_normalized`),
  ) %>%
  ungroup()

# MEAN NORMALIZATION
# Calculating individual channel intensities for PSMs
PSM_TMT126_mean_channel_signal <- mean(TMT_single_scan_df$`126_corrected`, na.rm = TRUE)
PSM_TMT127_mean_channel_signal <- mean(TMT_single_scan_df$`127_corrected`, na.rm = TRUE)
PSM_TMT128_mean_channel_signal <- mean(TMT_single_scan_df$`128_corrected`, na.rm = TRUE)
PSM_TMT129_mean_channel_signal <- mean(TMT_single_scan_df$`129_corrected`, na.rm = TRUE)
PSM_TMT130_mean_channel_signal <- mean(TMT_single_scan_df$`130_corrected`, na.rm = TRUE)
PSM_TMT131_mean_channel_signal <- mean(TMT_single_scan_df$`131_corrected`, na.rm = TRUE)

# Calculating channel normalization factors
mean_channel_signals <- c(PSM_TMT126_mean_channel_signal,
                          PSM_TMT127_mean_channel_signal,
                          PSM_TMT128_mean_channel_signal,
                          PSM_TMT129_mean_channel_signal,
                          PSM_TMT130_mean_channel_signal,
                          PSM_TMT131_mean_channel_signal)

max_mean_signal <- max(mean_channel_signals)

mean_channel_norm <- (mean_channel_signals / max_mean_signal)

# Normalizing individual channel intensities for PSMs
TMT_single_scan_df <- TMT_single_scan_df %>%
  rowwise() %>%
  mutate( `126_mean_normalized` = `126_corrected` / mean_channel_norm[1],
          `127_mean_normalized` = `127_corrected` / mean_channel_norm[2],
          `128_mean_normalized` = `128_corrected` / mean_channel_norm[3],
          `129_mean_normalized` = `129_corrected` / mean_channel_norm[4],
          `130_mean_normalized` = `130_corrected` / mean_channel_norm[5],
          `131_mean_normalized` = `131_corrected` / mean_channel_norm[6],
  ) %>%
  ungroup()

# log2 transform
TMT_single_scan_df <- TMT_single_scan_df %>%
  rowwise() %>%
  mutate( log2_mean_126 = log2(`126_mean_normalized`),
          log2_mean_127 = log2(`127_mean_normalized`),
          log2_mean_128 = log2(`128_mean_normalized`),
          log2_mean_129 = log2(`129_mean_normalized`),
          log2_mean_130 = log2(`130_mean_normalized`),
          log2_mean_131 = log2(`131_mean_normalized`),
  ) %>%
  ungroup()

# MEDIAN NORMALIZATION
# Calculating individual channel intensities for PSMs
PSM_TMT126_med_channel_signal <- median(TMT_single_scan_df$`126_corrected`, na.rm = TRUE)
PSM_TMT127_med_channel_signal <- median(TMT_single_scan_df$`127_corrected`, na.rm = TRUE)
PSM_TMT128_med_channel_signal <- median(TMT_single_scan_df$`128_corrected`, na.rm = TRUE)
PSM_TMT129_med_channel_signal <- median(TMT_single_scan_df$`129_corrected`, na.rm = TRUE)
PSM_TMT130_med_channel_signal <- median(TMT_single_scan_df$`130_corrected`, na.rm = TRUE)
PSM_TMT131_med_channel_signal <- median(TMT_single_scan_df$`131_corrected`, na.rm = TRUE)

# Calculating channel normalization factors
median_channel_signals <- c(PSM_TMT126_med_channel_signal,
                          PSM_TMT127_med_channel_signal,
                          PSM_TMT128_med_channel_signal,
                          PSM_TMT129_med_channel_signal,
                          PSM_TMT130_med_channel_signal,
                          PSM_TMT131_med_channel_signal)

max_med_signal <- max(median_channel_signals)

median_channel_norm <- (median_channel_signals / max_med_signal)

# Normalizing individual channel intensities for PSMs
TMT_single_scan_df <- TMT_single_scan_df %>%
  rowwise() %>%
  mutate( `126_median_normalized` = `126_corrected` / median_channel_norm[1],
          `127_median_normalized` = `127_corrected` / median_channel_norm[2],
          `128_median_normalized` = `128_corrected` / median_channel_norm[3],
          `129_median_normalized` = `129_corrected` / median_channel_norm[4],
          `130_median_normalized` = `130_corrected` / median_channel_norm[5],
          `131_median_normalized` = `131_corrected` / median_channel_norm[6],
  ) %>%
  ungroup()

# log2 transform
TMT_single_scan_df <- TMT_single_scan_df %>%
  rowwise() %>%
  mutate( log2_median_126 = log2(`126_median_normalized`),
          log2_median_127 = log2(`127_median_normalized`),
          log2_median_128 = log2(`128_median_normalized`),
          log2_median_129 = log2(`129_median_normalized`),
          log2_median_130 = log2(`130_median_normalized`),
          log2_median_131 = log2(`131_median_normalized`),
  ) %>%
  ungroup()

```

Comparing changes among samples following normalization

```{r}
# CUMULATIVE
# Using the t.test function in R to compare log2 transformed data
TMT_cumulative_ttest_df <- TMT_single_scan_df %>%
  group_by(scan_number)  %>%
  summarise(
    cum_t_test = list(tryCatch({
      t.test(c_across(c("log2_cum_126", "log2_cum_127", "log2_cum_128")),
             c_across(c("log2_cum_129", "log2_cum_130", "log2_cum_131")))
    }, error = function(e) {
      list(p.value = NA, statistic = NA, conf.int = c(NA, NA), estimate = c(NA, NA))
    }))
  ) # From t.test() default is "two.sided" (versus "less" or "greater")

# extracting the t-test results in their own dataframe because it didn't work after merging the ttest_df with the single_scan_df
TMT_cumulative_ttest_df <- TMT_cumulative_ttest_df %>%
  mutate(
    cum_p_value = map_dbl(cum_t_test, ~.x$p.value),
    cum_t_statistic = map_dbl(cum_t_test, ~.x$statistic),
    cum_conf_low = map_dbl(cum_t_test, ~.x$conf.int[1]),
    cum_conf_high = map_dbl(cum_t_test, ~.x$conf.int[2]),
    cum_estimate_mean_x = map_dbl(cum_t_test, ~.x$estimate['mean of x']),
    cum_estimate_mean_y = map_dbl(cum_t_test, ~.x$estimate['mean of y'])
  )  

# joining the ttest_df to the single_scan_df
TMT_single_scan_df <- TMT_single_scan_df %>%
  left_join(TMT_cumulative_ttest_df, by = "scan_number") %>%
  ungroup()

# MEAN
# Using the t.test function in R to compare log2 transformed data
TMT_mean_ttest_df <- TMT_single_scan_df %>%
  group_by(scan_number)  %>%
  summarise(
    mean_t_test = list(tryCatch({
      t.test(c_across(c("log2_mean_126", "log2_mean_127", "log2_mean_128")),
             c_across(c("log2_mean_129", "log2_mean_130", "log2_mean_131")))
    }, error = function(e) {
      list(p.value = NA, statistic = NA, conf.int = c(NA, NA), estimate = c(NA, NA))
    }))
  )

# extracting the t-test results in their own dataframe because it didn't work after merging the ttest_df with the single_scan_df
TMT_mean_ttest_df <- TMT_mean_ttest_df %>%
  mutate(
    mean_p_value = map_dbl(mean_t_test, ~.x$p.value),
    mean_t_statistic = map_dbl(mean_t_test, ~.x$statistic),
    mean_conf_low = map_dbl(mean_t_test, ~.x$conf.int[1]),
    mean_conf_high = map_dbl(mean_t_test, ~.x$conf.int[2]),
    mean_estimate_mean_x = map_dbl(mean_t_test, ~.x$estimate['mean of x']),
    mean_estimate_mean_y = map_dbl(mean_t_test, ~.x$estimate['mean of y'])
  )  

# joining the ttest_df to the single_scan_df
TMT_single_scan_df <- TMT_single_scan_df %>%
  left_join(TMT_mean_ttest_df, by = "scan_number") %>%
  ungroup()

# MEDIAN
# Using the t.test function in R to compare log2 transformed data
TMT_median_ttest_df <- TMT_single_scan_df %>%
  group_by(scan_number)  %>%
  summarise(
    median_t_test = list(tryCatch({
      t.test(c_across(c("log2_median_126", "log2_median_127", "log2_median_128")),
             c_across(c("log2_median_129", "log2_median_130", "log2_median_131")))
    }, error = function(e) {
      list(p.value = NA, statistic = NA, conf.int = c(NA, NA), estimate = c(NA, NA))
    }))
  )

# extracting the t-test results in their own dataframe because it didn't work after merging the ttest_df with the single_scan_df
TMT_median_ttest_df <- TMT_median_ttest_df %>%
  mutate(
    median_p_value = map_dbl(median_t_test, ~.x$p.value),
    median_t_statistic = map_dbl(median_t_test, ~.x$statistic),
    median_conf_low = map_dbl(median_t_test, ~.x$conf.int[1]),
    median_conf_high = map_dbl(median_t_test, ~.x$conf.int[2]),
    median_estimate_mean_x = map_dbl(median_t_test, ~.x$estimate['mean of x']),
    median_estimate_mean_y = map_dbl(median_t_test, ~.x$estimate['mean of y'])
  )  

# joining the ttest_df to the single_scan_df
TMT_single_scan_df <- TMT_single_scan_df %>%
  left_join(TMT_median_ttest_df, by = "scan_number") %>%
  ungroup()

# define columns for ttest_summary
column_order = c("scan_number", "126_cum_normalized", "127_cum_normalized", "128_cum_normalized", "129_cum_normalized", "130_cum_normalized", "131_cum_normalized", "log2_cum_126", "log2_cum_127", "log2_cum_128", "log2_cum_129", "log2_cum_130", "log2_cum_131", "126_mean_normalized", "127_mean_normalized", "128_mean_normalized", "129_mean_normalized", "130_mean_normalized", "131_mean_normalized", "log2_mean_126", "log2_mean_127", "log2_mean_128", "log2_mean_129", "log2_mean_130", "log2_mean_131", "126_median_normalized", "127_median_normalized", "128_median_normalized", "129_median_normalized", "130_median_normalized", "131_median_normalized", "log2_median_126", "log2_median_127", "log2_median_128", "log2_median_129", "log2_median_130", "log2_median_131", "cum_t_test", "cum_p_value", "cum_t_statistic", "cum_conf_low", "cum_conf_high", "cum_estimate_mean_x", "cum_estimate_mean_y", "mean_t_test", "mean_p_value", "mean_t_statistic", "mean_conf_low", "mean_conf_high", "mean_estimate_mean_x", "mean_estimate_mean_y", "median_t_test", "median_p_value", "median_t_statistic", "median_conf_low", "median_conf_high", "median_estimate_mean_x", "median_estimate_mean_y")

TMT_ttest_summary_df <- TMT_single_scan_df[, column_order]
    
# join results back to assigned_fractional_intensity_hits
assigned_fractional_intensity_hits <- assigned_fractional_intensity_hits %>%
  group_by(scan_number) %>%
  left_join(TMT_ttest_summary_df, by = "scan_number") %>%
  ungroup()

total_PSM_intensity <- sum(assigned_fractional_intensity_hits$total_intensity)

# Export RData version of analyzed hits
setwd(directory)
file_name <- paste0("experiment1_all_hits_assigned_fractional_intensity_hits_", save_name_RData_suffix)
save(assigned_fractional_intensity_hits, file = file_name)

# flatten list columns to strings for preservation when exporting to excel
assigned_hits_export <- assigned_fractional_intensity_hits %>%
  mutate(across(where(is.list), ~sapply(., toString)))

# Export excel .csv of analyzed hits
file_name <- paste0("experiment1_all_hits_assigned_fractional_intensity_hits_", save_name_csv_suffix)
write_excel_csv(assigned_hits_export, file = file_name)
```

Combining all single measurements of specific proteoforms and repeating t-test on aggregated data

```{r}
one_hit_vector <- as.vector(single_hit_scans$scan_number)

# CUMULATIVE
# Using the t.test function in R to compare log2 transformed data
TMT_cumulative_hit_grouped_ttest_df <- assigned_fractional_intensity_hits %>%
  filter(scan_number %in% one_hit_vector) %>%
  group_by(all_mods)  %>%
  summarise(
    cum_t_test = list(tryCatch({
      t.test(c_across(c("log2_cum_126", "log2_cum_127", "log2_cum_128")),
             c_across(c("log2_cum_129", "log2_cum_130", "log2_cum_131")))
    }, error = function(e) {
      list(p.value = NA, statistic = NA, conf.int = c(NA, NA), estimate = c(NA, NA))
    }))
  ) # From t.test() default is "two.sided" (versus "less" or "greater")

# extracting the t-test results in their own dataframe because it didn't work after merging the ttest_df with the single_scan_df
TMT_cumulative_hit_grouped_ttest_df <- TMT_cumulative_hit_grouped_ttest_df %>%
  mutate(
    cum_p_value = map_dbl(cum_t_test, ~.x$p.value),
    cum_t_statistic = map_dbl(cum_t_test, ~.x$statistic),
    cum_conf_low = map_dbl(cum_t_test, ~.x$conf.int[1]),
    cum_conf_high = map_dbl(cum_t_test, ~.x$conf.int[2]),
    cum_estimate_mean_x = map_dbl(cum_t_test, ~.x$estimate['mean of x']),
    cum_estimate_mean_y = map_dbl(cum_t_test, ~.x$estimate['mean of y'])
  )  

TMT_cumulative_hit_grouped_fc_df <- assigned_fractional_intensity_hits %>%
  filter(scan_number %in% one_hit_vector) %>%
  group_by(all_mods)  %>%
  mutate( `fc_cum` = (mean(c_across(c("log2_cum_129", "log2_cum_130", "log2_cum_131")), na.rm = TRUE) - mean(c_across(c("log2_cum_126", "log2_cum_127", "log2_cum_128")), na.rm = TRUE)))

TMT_cumulative_hit_grouped_fc_df <- TMT_cumulative_hit_grouped_fc_df %>%
  distinct(TMT_cumulative_hit_grouped_fc_df$all_mods, .keep_all = TRUE)

column_order = c("all_mods", "fc_cum")

TMT_cumulative_hit_grouped_fc_df <- TMT_cumulative_hit_grouped_fc_df[, column_order]

# joining the ttest_df to the single_scan_df
TMT_cumulative_hit_grouped_ttest_df <- TMT_cumulative_hit_grouped_ttest_df %>%
  group_by(all_mods)  %>%
  left_join(TMT_cumulative_hit_grouped_fc_df, by = "all_mods") %>%
  filter(!(is.na(cum_p_value))) %>%
  mutate(change = case_when(
    fc_cum < -0.6 & cum_p_value < 0.05 ~ "DOWN",
    abs(fc_cum) <= 0.6 | cum_p_value > 0.5 ~ "NO",
    fc_cum > 0.6 & cum_p_value < 0.05 ~ "UP",
    TRUE ~ "NO"
  )) %>%
  ungroup()

TMT_cumulative_hit_grouped_ttest_df <- TMT_cumulative_hit_grouped_ttest_df %>%
  mutate(label_cond = case_when(
    change %in% c("UP", "DOWN") & str_replace(all_mods, "Cterm36srt", "") == "" ~ "unmodified",
    change %in% c("UP", "DOWN") ~ str_replace(all_mods, "Cterm36srt", ""),
    TRUE ~ ""
  ))

# MEAN
# Using the t.test function in R to compare log2 transformed data
TMT_mean_hit_grouped_ttest_df <- assigned_fractional_intensity_hits %>%
  filter(scan_number %in% one_hit_vector) %>%
  group_by(all_mods)  %>%
  summarise(
    mean_t_test = list(tryCatch({
      t.test(c_across(c("log2_mean_126", "log2_mean_127", "log2_mean_128")),
             c_across(c("log2_mean_129", "log2_mean_130", "log2_mean_131")))
    }, error = function(e) {
      list(p.value = NA, statistic = NA, conf.int = c(NA, NA), estimate = c(NA, NA))
    }))
  ) # From t.test() default is "two.sided" (versus "less" or "greater")

# extracting the t-test results in their own dataframe because it didn't work after merging the ttest_df with the single_scan_df
TMT_mean_hit_grouped_ttest_df <- TMT_mean_hit_grouped_ttest_df %>%
  mutate(
    mean_p_value = map_dbl(mean_t_test, ~.x$p.value),
    mean_t_statistic = map_dbl(mean_t_test, ~.x$statistic),
    mean_conf_low = map_dbl(mean_t_test, ~.x$conf.int[1]),
    mean_conf_high = map_dbl(mean_t_test, ~.x$conf.int[2]),
    mean_estimate_mean_x = map_dbl(mean_t_test, ~.x$estimate['mean of x']),
    mean_estimate_mean_y = map_dbl(mean_t_test, ~.x$estimate['mean of y'])
  )  

TMT_mean_hit_grouped_fc_df <- assigned_fractional_intensity_hits %>%
  filter(scan_number %in% one_hit_vector) %>%
  group_by(all_mods)  %>%
  mutate( `fc_mean` = (mean(c_across(c("log2_mean_129", "log2_mean_130", "log2_mean_131")), na.rm = TRUE) - mean(c_across(c("log2_mean_126", "log2_mean_127", "log2_mean_128")), na.rm = TRUE)))

TMT_mean_hit_grouped_fc_df <- TMT_mean_hit_grouped_fc_df %>%
  distinct(TMT_mean_hit_grouped_fc_df$all_mods, .keep_all = TRUE)

column_order = c("all_mods", "fc_mean")

TMT_mean_hit_grouped_fc_df <- TMT_mean_hit_grouped_fc_df[, column_order]

# joining the ttest_df to the single_scan_df
TMT_mean_hit_grouped_ttest_df <- TMT_mean_hit_grouped_ttest_df %>%
  group_by(all_mods)  %>%
  left_join(TMT_mean_hit_grouped_fc_df, by = "all_mods") %>%
  filter(!(is.na(mean_p_value))) %>%
  mutate(change = case_when(
    fc_mean < -0.6 & mean_p_value < 0.05 ~ "DOWN",
    abs(fc_mean) <= 0.6 | mean_p_value > 0.5 ~ "NO",
    fc_mean > 0.6 & mean_p_value < 0.05 ~ "UP",
    TRUE ~ "NO"
  )) %>%
  ungroup()

TMT_mean_hit_grouped_ttest_df <- TMT_mean_hit_grouped_ttest_df %>%
  mutate(label_cond = case_when(
    change %in% c("UP", "DOWN") & str_replace(all_mods, "Cterm36srt", "") == "" ~ "unmodified",
    change %in% c("UP", "DOWN") ~ str_replace(all_mods, "Cterm36srt", ""),
    TRUE ~ ""
  ))

# MEDIAN
# Using the t.test function in R to compare log2 transformed data
TMT_median_hit_grouped_ttest_df <- assigned_fractional_intensity_hits %>%
  filter(scan_number %in% one_hit_vector) %>%
  group_by(all_mods)  %>%
  summarise(
    median_t_test = list(tryCatch({
      t.test(c_across(c("log2_median_126", "log2_median_127", "log2_median_128")),
             c_across(c("log2_median_129", "log2_median_130", "log2_median_131")))
    }, error = function(e) {
      list(p.value = NA, statistic = NA, conf.int = c(NA, NA), estimate = c(NA, NA))
    }))
  ) # From t.test() default is "two.sided" (versus "less" or "greater")

# extracting the t-test results in their own dataframe because it didn't work after merging the ttest_df with the single_scan_df
TMT_median_hit_grouped_ttest_df <- TMT_median_hit_grouped_ttest_df %>%
  mutate(
    median_p_value = map_dbl(median_t_test, ~.x$p.value),
    median_t_statistic = map_dbl(median_t_test, ~.x$statistic),
    median_conf_low = map_dbl(median_t_test, ~.x$conf.int[1]),
    median_conf_high = map_dbl(median_t_test, ~.x$conf.int[2]),
    median_estimate_mean_x = map_dbl(median_t_test, ~.x$estimate['mean of x']),
    median_estimate_mean_y = map_dbl(median_t_test, ~.x$estimate['mean of y'])
  )  

TMT_median_hit_grouped_fc_df <- assigned_fractional_intensity_hits %>%
  filter(scan_number %in% one_hit_vector) %>%
  group_by(all_mods)  %>%
  mutate( `fc_median` = (mean(c_across(c("log2_median_129", "log2_median_130", "log2_median_131")), na.rm = TRUE) - mean(c_across(c("log2_median_126", "log2_median_127", "log2_median_128")), na.rm = TRUE)))

TMT_median_hit_grouped_fc_df <- TMT_median_hit_grouped_fc_df %>%
  distinct(TMT_median_hit_grouped_fc_df$all_mods, .keep_all = TRUE)

column_order = c("all_mods", "fc_median")

TMT_median_hit_grouped_fc_df <- TMT_median_hit_grouped_fc_df[, column_order]

# joining the ttest_df to the single_scan_df
TMT_median_hit_grouped_ttest_df <- TMT_median_hit_grouped_ttest_df %>%
  group_by(all_mods)  %>%
  left_join(TMT_median_hit_grouped_fc_df, by = "all_mods") %>%
  filter(!(is.na(median_p_value))) %>%
  mutate(change = case_when(
    fc_median < -0.6 & median_p_value < 0.05 ~ "DOWN",
    abs(fc_median) <= 0.6 | median_p_value > 0.5 ~ "NO",
    fc_median > 0.6 & median_p_value < 0.05 ~ "UP",
    TRUE ~ "NO"
  )) %>%
  ungroup()

TMT_median_hit_grouped_ttest_df <- TMT_median_hit_grouped_ttest_df %>%
  mutate(label_cond = case_when(
    change %in% c("UP", "DOWN") & str_replace(all_mods, "Cterm36srt", "") == "" ~ "unmodified",
    change %in% c("UP", "DOWN") ~ str_replace(all_mods, "Cterm36srt", ""),
    TRUE ~ ""
  ))

# Export RData version of t-test analysis
setwd(directory)
file_name <- paste0("experiment1_all_hits_cumulative_ttest_", save_name_RData_suffix)
save(TMT_cumulative_hit_grouped_ttest_df, file = file_name)
file_name <- paste0("experiment1_all_hits_mean_ttest_", save_name_RData_suffix)
save(TMT_mean_hit_grouped_ttest_df, file = file_name)
file_name <- paste0("experiment1_all_hits_median_ttest_", save_name_RData_suffix)
save(TMT_median_hit_grouped_ttest_df, file = file_name)
```

Quantifying proteoforms for further description of species with significant fold changes

```{r}
proteoform_intensity_df <- assigned_fractional_intensity_hits %>%
  group_by(all_mods) %>%
  mutate( proteoform_intensity = sum(total_intensity)) %>%
  ungroup()

column_order <- c("scan_number", "PID", "all_mods", "total_intensity", "proteoform_intensity")

proteoform_intensity_df <- proteoform_intensity_df[, column_order]

proteoform_intensity_df <- proteoform_intensity_df %>%
  filter(proteoform_intensity_df$all_mods %in% unique(proteoform_intensity_df$all_mods))

# CUMULATIVE
proteoform_cumulative_vector <- as.vector(TMT_cumulative_hit_grouped_fc_df$all_mods)

proteoform_intensity_cumulative_df <- data.frame(proteoform_cumulative_vector)

names(proteoform_intensity_cumulative_df) <- c("all_mods")

proteoform_intensity_cumulative_df<- proteoform_intensity_cumulative_df %>%
  left_join(proteoform_intensity_df, by = "all_mods") %>%
  distinct(all_mods, .keep_all = TRUE) %>%
  left_join(TMT_cumulative_hit_grouped_ttest_df, by = "all_mods") %>%
  rowwise() %>%
  mutate( ctrl_int = (1 / (2^(fc_cum) + 1)) * proteoform_intensity,
          var_int = ((2^(fc_cum)) / (2^(fc_cum) + 1)) * proteoform_intensity
  ) %>%
  mutate( ctrl_percent = ctrl_int / total_PSM_intensity,
          var_percent = var_int / total_PSM_intensity) %>%
  mutate("var_abundance" = var_percent) %>%
  ungroup() %>%
  filter(!(is.na(cum_p_value))) 

# confirmed that summation is working correctly (see below)
# sum(proteoform_intensity_cumulative_df$total_intensity[1:109])

# MEAN
proteoform_mean_vector <- as.vector(TMT_mean_hit_grouped_fc_df$all_mods)

proteoform_intensity_mean_df <- data.frame(proteoform_mean_vector)

names(proteoform_intensity_mean_df) <- c("all_mods")

proteoform_intensity_mean_df<- proteoform_intensity_mean_df %>%
  left_join(proteoform_intensity_df, by = "all_mods") %>%
  distinct(all_mods, .keep_all = TRUE) %>%
  left_join(TMT_mean_hit_grouped_ttest_df, by = "all_mods") %>%
  rowwise() %>%
  mutate( ctrl_int = (1 / (2^(fc_mean) + 1)) * proteoform_intensity,
          var_int = ((2^(fc_mean)) / (2^(fc_mean) + 1)) * proteoform_intensity
  ) %>%
  mutate( ctrl_percent = ctrl_int / total_PSM_intensity,
          var_percent = var_int / total_PSM_intensity) %>%
  mutate("var_abundance" = var_percent) %>%
  ungroup() %>%
  filter(!(is.na(mean_p_value)))

# MEDIAN
proteoform_median_vector <- as.vector(TMT_median_hit_grouped_fc_df$all_mods)

proteoform_intensity_median_df <- data.frame(proteoform_median_vector)

names(proteoform_intensity_median_df) <- c("all_mods")

proteoform_intensity_median_df<- proteoform_intensity_median_df %>%
  left_join(proteoform_intensity_df, by = "all_mods") %>%
  distinct(all_mods, .keep_all = TRUE) %>%
  left_join(TMT_median_hit_grouped_ttest_df, by = "all_mods") %>%
  rowwise() %>%
  mutate( ctrl_int = (1 / (2^(fc_median) + 1)) * proteoform_intensity,
          var_int = ((2^(fc_median)) / (2^(fc_median) + 1)) * proteoform_intensity
  ) %>%
  mutate( ctrl_percent = ctrl_int / total_PSM_intensity,
          var_percent = var_int / total_PSM_intensity) %>%
  mutate("var_abundance" = var_percent) %>%
  ungroup()%>%
  filter(!(is.na(median_p_value)))
```

Plotting

```{r}
# CUMULATIVE
experiment1_cumulative_norm_volcano_plot <- ggplot(data = TMT_cumulative_hit_grouped_ttest_df, aes(x = fc_cum, y = -log10(cum_p_value), col = change, label = label_cond)) +
  geom_vline(xintercept = c(-0.6, 0.6), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
  geom_point(size = 2) +
  scale_color_manual(values = c("#00AFBB", "grey", "#bb0c00"), # to set the colours of our variable
                       labels = c("Decreased", "No significant change", "Increased")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  coord_cartesian(ylim = c(0, 50), xlim = c(-3, 3)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = 'Proteoform Change', #legend_title
       x = expression("log"[2]*"FC"), y = expression("-log"[10]*"p-value")) +
  scale_x_continuous(breaks = seq(-10, 10, 2)) + # to customise the breaks in the x axis
  ggtitle('Proteoforms changed on experiment1 treatment - cumulative normalization') + # Plot title
  geom_text_repel(
    size = 3, 
    max.overlaps = Inf,
    #nudge_x = 0.05, 
    nudge_y = 0.15,
    point.padding = unit(0.4, "lines"),  # adjust based on your specific plot's scale
    box.padding = unit(0.4, "lines")
  ) # To show all labels

# MEAN
experiment1_mean_norm_volcano_plot <- ggplot(data = TMT_mean_hit_grouped_ttest_df, aes(x = fc_mean, y = -log10(mean_p_value), col = change, label = label_cond)) +
  geom_vline(xintercept = c(-0.6, 0.6), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
  geom_point(size = 2) +
  scale_color_manual(values = c("#00AFBB", "grey", "#bb0c00"), # to set the colours of our variable
                       labels = c("Decreased", "No significant change", "Increased")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  coord_cartesian(ylim = c(0, 50), xlim = c(-3, 3)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = 'Proteoform Change', #legend_title
       x = expression("log"[2]*"FC"), y = expression("-log"[10]*"p-value")) +
  scale_x_continuous(breaks = seq(-10, 10, 2)) + # to customise the breaks in the x axis
  ggtitle('Proteoforms changed on experiment1 treatment - mean normalization') + # Plot title
  geom_text_repel(
    size = 3, 
    max.overlaps = Inf,
    #nudge_x = 0.05, 
    nudge_y = 0.15,
    point.padding = unit(0.4, "lines"),  # adjust based on your specific plot's scale
    box.padding = unit(0.4, "lines")
  ) # To show all labels

# CUMULATIVE
experiment1_median_norm_volcano_plot <- ggplot(data = TMT_median_hit_grouped_ttest_df, aes(x = fc_median, y = -log10(median_p_value), col = change, label = label_cond)) +
  geom_vline(xintercept = c(-0.6, 0.6), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
  geom_point(size = 2) +
  scale_color_manual(values = c("#00AFBB", "grey", "#bb0c00"), # to set the colours of our variable
                       labels = c("Decreased", "No significant change", "Increased")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  coord_cartesian(ylim = c(0, 50), xlim = c(-3, 3)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = 'Proteoform Change', #legend_title
       x = expression("log"[2]*"FC"), y = expression("-log"[10]*"p-value")) +
  scale_x_continuous(breaks = seq(-10, 10, 2)) + # to customise the breaks in the x axis
  ggtitle('Proteoforms changed on experiment1 treatment - median normalization') + # Plot title
  geom_text_repel(
    size = 3, 
    max.overlaps = Inf,
    #nudge_x = 0.05, 
    nudge_y = 0.15,
    point.padding = unit(0.4, "lines"),  # adjust based on your specific plot's scale
    box.padding = unit(0.4, "lines")
  ) # To show all labels

# Export RData version of t-test analysis
setwd(directory)
file_name <- paste0("experiment1_all_hits_cumulative_volcano_plot_", save_name_RData_suffix)
save(experiment1_cumulative_norm_volcano_plot, file = file_name)
file_name <- paste0("experiment1_all_hits_mean_volcano_plot_", save_name_RData_suffix)
save(experiment1_mean_norm_volcano_plot, file = file_name)
file_name <- paste0("experiment1_all_hits_median_volcano_plot_", save_name_RData_suffix)
save(experiment1_median_norm_volcano_plot, file = file_name)
```

Printing tables for comparison

```{r}
proteoform_df <- assigned_fractional_intensity_hits %>%
  group_by(all_mods) %>%
  mutate( proteoform_intensity = sum(total_intensity)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate( fractional_intensity = proteoform_intensity / total_PSM_intensity) %>%
  ungroup() %>%
  distinct(all_mods, .keep_all = TRUE)

column_order <- c("all_mods", "proteoform_intensity", "fractional_intensity")  
  
proteoform_df <- proteoform_df[, column_order]

names(proteoform_df) <- c("all_mods", "proteoform_intensity", "percent_abundance")

column_order <- c("all_mods", "cum_p_value", "fc_cum", "ctrl_percent", "var_percent")  
cumulative_df <- proteoform_intensity_cumulative_df
  
cumulative_df <- cumulative_df[, column_order]
    
names(cumulative_df) <- c("all_mods", "cumulative_norm_p_value", "cumulative_log2-fold_change", "cumulative_ctrl_percent", "cumulative_var_percent")

column_order <- c("all_mods", "mean_p_value", "fc_mean", "ctrl_percent", "var_percent")  

mean_df <- proteoform_intensity_mean_df
  
mean_df <- mean_df[, column_order]
    
names(mean_df) <- c("all_mods", "mean_norm_p_value", "mean_log2-fold_change", "mean_ctrl_percent", "mean_var_percent")

column_order <- c("all_mods", "median_p_value", "fc_median", "ctrl_percent", "var_percent")  

median_df <- proteoform_intensity_median_df
  
median_df <- median_df[, column_order]
    
names(median_df) <- c("all_mods", "median_norm_p_value", "median_log2-fold_change", "median_ctrl_percent", "median_var_percent")

proteoform_df <- proteoform_df %>%
  left_join(cumulative_df, by = "all_mods") %>%
  left_join(mean_df, by = "all_mods") %>%
  left_join(median_df, by = "all_mods") %>%
  mutate(all_mods = sub(" Cterm36srt", "", all_mods)) %>%
  mutate(all_mods = sub("Cterm36srt", "unmodified", all_mods))

setwd(directory)
file_name <- paste0("experiment1_proteoform_df_", save_name_csv_suffix)
write_excel_csv(proteoform_df, file = file_name)

file_name <- paste0("experiment1_proteoform_df_", save_name_RData_suffix)
save(proteoform_df, file = file_name)
```



