---
title: "Ion Calculator"
output: html_notebook
---
This was incorporated into, and updated in Disambiguator.Rmd

Libraries & working directory

```{r}
directory <- c("C:/TPP/data/test")
setwd(directory)

library(Spectra)
library(mzR)
library(stringr)
library(readxl)
library(dplyr)
library(xml2)

# Replace 'path_to_your_excel_file.xlsx' with the actual path to your Excel file
# excel file should contain scan numbers and peptide sequences following the unusual format described below
# excel_data <- read_excel('path_to_your_excel_file.xlsx')
# Enter the path to your proteomics data. This should include your mzML.gz, .raw, pep.xml and .PIN files

my_pep_xml <- c("20240215_EBZ_Cole_20ng_293f_MS275vDMSO_1to1_6plex_2-15grad.pep.xml")

# Load your XML data
peptide_data <- read_xml(my_pep_xml)

# Replace 'd1' and its URI with the correct namespace prefix and URI if your XML uses namespaces
ns <- xml_ns_rename(xml_ns(peptide_data), c(d1 = "d1"))
```

Step 1: Define Masses of Amino Acids and Atoms
Start by defining the monoisotopic masses for amino acids and the relevant atomic masses. This will serve as the foundation for all subsequent calculations.

```{r}
# Monoisotopic masses of amino acids
amino_acid_masses <- c(
  A = 71.03711, R = 156.10111, N = 114.04293, D = 115.02694,
  C = 103.00919, E = 129.04259, Q = 128.05858, G = 57.02146,
  H = 137.05891, I = 113.08406, L = 113.08406, K = 128.09496,
  M = 131.04049, F = 147.06841, P = 97.05276, S = 87.03203,
  T = 101.04768, W = 186.07931, Y = 163.06333, V = 99.06841
)

# Atomic masses for elements (Monoisotopic)
atomic_masses <- c(
  C = 12.0, c = 13.0033548378, H = 1.007825, N = 14.003074, n = 15.0001088982, O = 15.994915, P = 30.973762, S = 31.972071
)

# PTMs
amino_acid_masses["a"] <- amino_acid_masses["K"] + (atomic_masses["C"] * 2 + atomic_masses["H"] * 2 + atomic_masses["O"]) # Kac
amino_acid_masses["1"] <- amino_acid_masses["K"] + (atomic_masses["C"] * 1 + atomic_masses["H"] * 2) # Kme1
amino_acid_masses["2"] <- amino_acid_masses["K"] + (atomic_masses["C"] * 2 + atomic_masses["H"] * 4) # Kme2
amino_acid_masses["3"] <- amino_acid_masses["K"] + (atomic_masses["C"] * 3 + atomic_masses["H"] * 6) # Kme3
amino_acid_masses["m"] <- amino_acid_masses["R"] + (atomic_masses["C"] * 1 + atomic_masses["H"] * 2) # Rme1
amino_acid_masses["d"] <- amino_acid_masses["R"] + (atomic_masses["C"] * 2 + atomic_masses["H"] * 4) # Rme2
amino_acid_masses["s"] <- amino_acid_masses["S"] + (atomic_masses["P"] + atomic_masses["H"] * 2 + atomic_masses["O"] * 3) # Sphos
amino_acid_masses["t"] <- amino_acid_masses["S"] + (atomic_masses["P"] + atomic_masses["H"] * 2 + atomic_masses["O"] * 3) # Sphos
amino_acid_masses["p"] <- amino_acid_masses["K"] + (atomic_masses["C"] * 3 + atomic_masses["H"] * 5 + atomic_masses["O"]) # Kprop
amino_acid_masses["b"] <- amino_acid_masses["K"] + (atomic_masses["C"] * 4 + atomic_masses["H"] * 6 + atomic_masses["O"] * 2) # Kbhb
amino_acid_masses["c"] <- amino_acid_masses["G"] + amino_acid_masses["K"] + (atomic_masses["C"] * 11 + atomic_masses["c"] * 4 + atomic_masses["H"] * (28 - 1) + atomic_masses["O"] * (3 - 1) + atomic_masses["N"] * 4 + atomic_masses["n"] * 1) # c-term tmt mod

print(amino_acid_masses)

# Adjustments for ion types
ion_type_adjustments <- list(
  b = 0, # [N]+[M]-H  ; N = mass of neutral N terminal group
  y = atomic_masses["H"] * (2) + atomic_masses["O"], # Neutral mass = [C]+[M]+H ; C = mass of neutral C-terminal group (OH)
  c = atomic_masses["H"] + (atomic_masses["H"] * (2) + atomic_masses["N"]), # [N]+[M]+NH2  ; N = mass of neutral N terminal group
  z = atomic_masses["O"] - atomic_masses["N"] # Neutral mass = [C]+[M]-NH2 ; C = mass of neutral C-terminal group (OH)
	# We're really printing Z-dot ions so we add an H to make it OH+[M]-NH2 +H = [M]+O-N
)
```

Step 1: Define Necessary Constants and Functions
First, ensure you have defined all the necessary constants, including amino acid and atomic masses, as well as functions for calculating peptide and ion masses.

```{r}
calculate_fragment_mass_from_sequence <- function(sequence) {
  mass <- sum(sapply(strsplit(sequence, "")[[1]], function(aa) {
    if (aa %in% names(amino_acid_masses)) {
      return(amino_acid_masses[[aa]])
    } else {
      stop(paste("Undefined amino acid or modification:", aa))
    }
  }))
  return(mass)
}

# Function to calculate m/z for ions
calculate_ion_mz <- function(peptide_mass, charge) {
  # Adjust mass for ion type and add protons for charge
  mz <- (peptide_mass + atomic_masses["H"] * charge) / charge
  return(mz)
}

generate_ions_for_sequence <- function(sequence) {
  ions_list <- list()
  
  sequence_length <- nchar(sequence)
  for (charge in 1:6) {
    b_ions <- numeric(sequence_length - 1)
    y_ions <- numeric(sequence_length - 1)
    c_ions <- numeric(sequence_length - 1)
    z_ions <- numeric(sequence_length - 1)
    
    for (i in 1:(sequence_length - 1)) {
      b_seq <- substr(sequence, 1, i)
      y_seq <- substr(sequence, i + 1, sequence_length)
      
      # Calculate the mass of the b, y, c, z ion sequences
      b_mass <- calculate_fragment_mass_from_sequence(b_seq) + ion_type_adjustments$b
      y_mass <- calculate_fragment_mass_from_sequence(y_seq) + ion_type_adjustments$y
      c_mass <- calculate_fragment_mass_from_sequence(b_seq) + ion_type_adjustments$c
      z_mass <- calculate_fragment_mass_from_sequence(y_seq) + ion_type_adjustments$z
      
      # Calculate the m/z for b, y, c, z ions. Note the adjustment for ion type is now applied in the mass calculation
      b_ions[i] <- calculate_ion_mz(b_mass, charge)
      y_ions[i] <- calculate_ion_mz(y_mass, charge)
      c_ions[i] <- calculate_ion_mz(c_mass, charge)
      z_ions[i] <- calculate_ion_mz(z_mass, charge)
    }
    
    # Populate ions_list with b, y, c, and z ions (c and z ions omitted for brevity)
    ions_list[[paste("b_ions_charge", charge)]] <- b_ions
    ions_list[[paste("y_ions_charge", charge)]] <- y_ions
    ions_list[[paste("c_ions_charge", charge)]] <- c_ions
    ions_list[[paste("z_ions_charge", charge)]] <- z_ions
  }
  
  return(ions_list)
}
```

Step 2: Generate Ions and Store in Vectors
You will need to iterate over the peptide sequence to generate b and y ions (and similarly c and z ions if desired) for each charge state from +1 to +6. Here, we focus on generating b and y ions as examples.

```{r}
# Example usage for a peptide sequence
peptide_sequence <- "ARTKQTARaSTGG2APRKQLATKAAR2SAPATGGc"
ions <- as.data.frame(generate_ions_for_sequence(peptide_sequence))
ions_df <- ions_df %>%
  mutate(across(starts_with("y_ions_charge"), rev)) %>%
  mutate(across(starts_with("z_ions_charge"), rev))
```

Step 3: Compare Against Mass Spectrum
Assuming you have a numerical list from a mass spectrum, you can compare the generated ion vectors to this list. Here’s a simple way to find matches within a specified tolerance:

```{r}
# Define fn to calculate tolerance
calculate_tolerance <- function(mz, ppm) {
  return(as.numeric(mz) * ppm / 1e6)
}

find_matches_with_intensity_df <- function(ions_df, scan_data, ppm_tolerance, sequence) {
  matches <- list()
  sequence_length <- nchar(sequence)
  half_sequence_length <- floor(sequence_length * 0.5)
  
  # Identify ion type columns in the dataframe
  ion_columns <- names(ions_df)
  
  for (ion_column in ion_columns) {
    # Determine the charge state based on the column name
    charge_state <- as.numeric(gsub(".*charge\\.(\\d+)$", "\\1", ion_column))
    ion_vector <- ions_df[[ion_column]]
    
    for (i in seq_along(ion_vector)) {
      # Skip condition for +1 charge state ions if their position is greater than half the sequence length
      if (charge_state == 1 && i > half_sequence_length) {
        next
      }

      # Additional conditions for other charge states if needed
      # Example for +3 charge state (you can add logic inside the block)
      else if (charge_state == 3 && i < 2) {
        # Specific logic for +3 charge state ions
      }
      
      # Condition for +4 charge state ions
      else if (charge_state == 4 && i < 3) {
      # Apply specific logic for +2 charge state ions
      # You can add more specific conditions for +2 charge state ions here
      }
      
      # Condition for +5 charge state ions
      else if (charge_state == 5 && i < 4) {
      # Apply specific logic for +2 charge state ions
      # You can add more specific conditions for +2 charge state ions here
      }
      
      # Condition for +5 charge state ions
      else if (charge_state == 6 && i < 4) {
      # Apply specific logic for +2 charge state ions
      # You can add more specific conditions for +2 charge state ions here
      }
      
      ion_mass <- ion_vector[i]
      ion_tolerance <- calculate_tolerance(ion_mass, ppm_tolerance)
      
      # Find indices of matching m/z values within tolerance
      matching_indices <- which(abs(scan_data[, "mz"] - ion_mass) <= ion_tolerance)
      
      if (length(matching_indices) > 0) {
        # For each matching index, capture the intensity
        intensities <- scan_data[matching_indices, "intensity"]
        # Construct a key for this match, including the column name and ion position
        match_key <- paste(ion_column, "pos", i, round(ion_mass, 4), sep = "_")
        # Store matches with their intensities
        matches[[match_key]] <- list(mz = scan_data[matching_indices, "mz"], intensity = intensities)
      }
    }
  }
  
  return(matches)
}

```

Key Points:
Charge State Detection: The function now includes logic to determine the charge state from the column name. This uses a regular expression with gsub() to extract the charge state number from the column name pattern like "y_ions_charge.1". Adjust the pattern as necessary to match your dataframe's column naming convention.

Biochemical Rule Application: Before proceeding with the match finding for each ion, it checks if the ion is in the +1 charge state and if its position is greater than half the length of the sequence. If both conditions are met, the next statement skips the rest of the current iteration, effectively filtering out these ions from the match list.

Sequence Length Consideration: The function now requires the sequence parameter to calculate the sequence length and thereby apply the biochemical rule accurately.

This approach allows you to integrate specific biochemical rules into the process of identifying and retaining meaningful matches, enhancing the biological relevance of your analysis results.
---
Extracting mzML spectrum data with Spectra and mzR
https://rformassspectrometry.r-universe.dev/articles/Spectra/Spectra.html

```{r}
setwd("C:/TPP/data/293f_2to15_DMSO_MS275_1to1_6plex")
mz <- openMSfile("20240215_EBZ_Cole_20ng_293f_MS275vDMSO_1to1_6plex_2-15grad.mzML.gz")
p <- peaks(mz) ## extract all peak information
scan2012 <- peaks(mz, scan=2012)

scan2012 <- as.data.frame(scan2012)

ppm_tolerance <- 20 # Define your ppm tolerance
matches_with_intensity <- find_matches_with_intensity(ions_df, scan2012, ppm_tolerance)

# Initialize lists to hold extracted data
ion_type_list <- c()
ion_position_list <- c()
ion_mz_list <- c()
match_mz_list <- c()
match_intensity_list <- c()

# Iterate through matches to extract data
for (match_key in names(matches_with_intensity)) {
  # Correctly split the match_key to extract details
  details <- unlist(strsplit(match_key, "_pos_")) # Split at "_pos_"
  ion_type_and_charge <- details[1] # This is the ion type and charge
  pos_and_mz <- unlist(strsplit(details[2], "_")) # Further split to separate position from m/z
  ion_position <- as.numeric(pos_and_mz[1]) # This is the position
  ion_mz <- as.numeric(pos_and_mz[2]) # This is the ion m/z
  
  # Extract matched m/z and intensity from each match
  for (i in seq_along(matches_with_intensity[[match_key]]$mz)) {
    ion_type_list <- c(ion_type_list, ion_type_and_charge)
    ion_position_list <- c(ion_position_list, ion_position)
    ion_mz_list <- c(ion_mz_list, ion_mz)
    match_mz_list <- c(match_mz_list, matches_with_intensity[[match_key]]$mz[i])
    match_intensity_list <- c(match_intensity_list, matches_with_intensity[[match_key]]$intensity[i])
  }
}

# Combine the lists into a dataframe
matches_df <- data.frame(
  IonType = ion_type_list,
  IonPosition = ion_position_list,
  IonMZ = ion_mz_list,
  MatchMZ = match_mz_list,
  MatchIntensity = match_intensity_list
)
```

Instructions for iterating over data set

```{r}
# Assuming 'mz' is already opened using openMSfile
results <- list()

for (i in 1:nrow(excel_data)) {
  scan_number <- excel_data$ScanNumber[i]
  peptide_sequence <- excel_data$PeptideSequence[i]
  
  # Generate scan data frame
  scan_data <- as.data.frame(peaks(mz, scan = scan_number))
  
  # Generate ions for peptide sequence
  ions_df <- as.data.frame(generate_ions_for_sequence(peptide_sequence))
  ions_df <- ions_df %>%
  mutate(across(starts_with("y_ions_charge"), rev)) %>%
  mutate(across(starts_with("z_ions_charge"), rev))
  
  # Find matches
  ppm_tolerance <- 20 # Define your ppm tolerance
  matches_with_intensity <- find_matches_with_intensity(ions_df, scan_data, ppm_tolerance)
  
  # Iterate through matches to extract data
  for (match_key in names(matches_with_intensity)) {
    # Correctly split the match_key to extract details
    details <- unlist(strsplit(match_key, "_pos_")) # Split at "_pos_"
    ion_type_and_charge <- details[1] # This is the ion type and charge
    pos_and_mz <- unlist(strsplit(details[2], "_")) # Further split to separate position from m/z
    ion_position <- as.numeric(pos_and_mz[1]) # This is the position
    ion_mz <- as.numeric(pos_and_mz[2]) # This is the ion m/z
  
    # Extract matched m/z and intensity from each match
    for (i in seq_along(matches_with_intensity[[match_key]]$mz)) {
      ion_type_list <- c(ion_type_list, ion_type_and_charge)
      ion_position_list <- c(ion_position_list, ion_position)
      ion_mz_list <- c(ion_mz_list, ion_mz)
      match_mz_list <- c(match_mz_list, matches_with_intensity[[match_key]]$mz[i])
      match_intensity_list <- c(match_intensity_list, matches_with_intensity[[match_key]]$intensity[i])
    }
  }

  # Combine the lists into a dataframe
  matches_df <- data.frame(
    IonType = ion_type_list,
    IonPosition = ion_position_list,
    IonMZ = ion_mz_list,
    MatchMZ = match_mz_list,
    MatchIntensity = match_intensity_list
  )
  
  # Store results
  results[[paste("Scan", scan_number)]] <- matches_df
  
  # Optional: Clear memory of temporary objects
  rm(scan_data, ions, matches_with_intensity, matches_df)
}

# Process 'results' as needed, e.g., save to files or further analysis
```



